### 2024-11-21 paper/paper.qmd
update paper with model and data part

---
title: "Examining the impact Overturn of Roe v. Wade: Banning of Abortion on Infant Mortality Rates in the United States Using a Difference-in-Differences Approach"
subtitle: "The infant death rate increased by 0.285 deaths per 1,000 live births in states that became abortion-banned, compared to states where abortion remained legal."
author: 
  - Diana Shen
thanks: "Code and data are available at: [https://github.com/DianaShen1224/Relationship-between-infant-mortality-rate-and-prohibited-abortion)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
editor: 
  markdown: 
    wrap: 72
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(arrow)
library(knitr)
```

```{r}
#| include: false
#| warning: false
#| message: false
analysis_data<-read_parquet(here::here("data", "02-analysis_data", "analysis_data.parquet"))
```

# Introduction

The overturning of Roe v. Wade on June 24, 2022, through the Supreme
Court’s decision in Dobbs v. Jackson Women’s Health Organization, ended
nearly five decades of constitutional protection for abortion rights.
This landmark ruling shifted authority over abortion regulations to
individual states, creating a patchwork of laws nationwide. The decision
reignited debates about reproductive rights and public health,
highlighting disparities in access to abortion services and their
effects on health outcomes, especially in states with restrictive
policies.

Banning abortion usually has been linked to a rise in maternal mortality
rates, with Texas experiencing a 56% increase from 2019 to 2022
[@nbcnews]. However, restrictive abortion laws can also impact infant
mortality, as they often lead to more unintended or high-risk
pregnancies among disadvantaged populations with limited access to
healthcare. These pregnancies are associated with higher rates of
preterm births, low birth weights, and delivery complications, all of
which elevate infant mortality rates. Even before Roe v. Wade was
overturned, researchers highlighted the impacts of restrictive abortion
laws on mortality rates. @Burdick2024 found that states with 11–12
restrictive laws during 2014–2018 had a 16% higher infant mortality
rate, particularly in counties with higher proportions of Black
populations, inadequate prenatal care, and maternal smoking. Similarly,
@Harper2023 showed that restrictive laws are linked to higher maternal,
fetal, and infant mortality, states with moderate abortion legislation
had significantly lower maternal mortality (25.79 per 100,000 live
births) and infant mortality (20.56 per 1,000 live births) compared to
restrictive states, underscoring the health benefits of less restrictive
policies.

The estimand of our analysis is the change in the difference in infant
death rates between states that banned abortion and states where
abortion remained legal after June 24, 2022. The object of the
estimation is to estimate the change in the difference in the infant
death rate between the treatment and control groups based on
observational data in the United States.

Instead of examining the impact of abortion laws over broader time
periods like prior research, our study focuses on the effect of abortion
bans on infant mortality rates in the United States specifically from
2021 to 2022, with an emphasis on changes following the 2022 Dobbs
decision. Using a Difference-in-Differences (DID) approach, we assess
the interaction between abortion legality in states and the timing of
the injunction implemented to understand its influence on infant health
outcomes. The study utilizes observational data from the Centers for
Disease Control and Prevention’s WONDER database, capturing infant
mortality trends across states with varying levels of abortion
restrictiveness. Key predictors include the mother’s state of residence,
race, age, and the year of infant death, enabling a nuanced examination
of the effects of restrictive abortion laws on public health.

Our Results show that the overturning of Roe v. Wade and the
implementation of abortion bans are linked to a rise in infant mortality
rates in states where abortion became illegal. Through a
Difference-in-Differences (DID) approach, we estimate that states
enforcing abortion bans after the injunction saw an increase of 0.285
deaths per 1,000 live births compared to states where abortion remained
legal. Furthermore, we find that maternal age and race are key
determinants of infant mortality, with older maternal age groups
associated with lower mortality rates and Black mothers experiencing
significantly higher rates. These results emphasize the public health
consequences of restrictive abortion laws and the urgent need for
policies that mitigate these disparities.

This study is important as we highlight the public health implications
of abortion bans, particularly their role in increasing infant mortality
rates. By focusing on the 2022 Dobbs decision, we provide timely
evidence on how restrictive abortion policies disproportionately affect
vulnerable populations, emphasizing the need for data-driven policies to
mitigate these adverse outcomes.

The structure of this paper is organized as follows: @sec-data provides
details on the data sources and variables used in our analysis.
@sec-model explains the modeling approach, including the assumptions and
specifications of our linear regression framework. In @sec-result, we
present our findings, emphasizing the key predictors of Harris's
support. Finally, @sec-discussion explores the implications of our
results and suggests potential directions for future research. @sec-a
provides external data detail, @sec-b provides model detail, @sec-c
provides an datasheet, and @sec-d provides an detailed introduction of
surveys, sampling and observational data in our paper.

# Data {#sec-data}

## Overview

We conduct our infant death rate observational data analysis using the R
programming language [@citeR]. Our dataset, obtained from CDC
Wonder[@cdc_wonder_2024], based on data as of December 2022, provides a
detailed overview of the Infant Death Rate across the USA. CDC WONDER
(Wide-ranging ONline Data for Epidemiologic Research) is an online
system providing public health professionals and the public with access
to a wide range of the Centers for Disease Control and Prevention (CDC)
epidemiologic data and resources. Following the guidelines outlined in
@tellingstories, we analyze the impact of abortion injunctions on infant
death rates using the Difference-in-Difference method. This approach
utilizes the interaction between the legality of abortion in the
mother’s state of residence and the timing of the infant’s death as the
predictor variable. Additionally, we include control variables such as
infant year of death, the mother’s state of residence, race, and age to
account for potential confounding factors.

## Measurement

The process of translating real-world events into our dataset requires a
systematic approach to measurement and data gathering. In this research,
we analyze infant mortality rates in the United States to understand the
impact of abortion policies and demographic factors. The CDC’s Linked
Birth/Infant Death Records database provides comprehensive data,
including maternal and infant characteristics, to support this study.
Variables such as maternal age, race, education, birth weight,
gestational age, and prenatal care are included to ensure a detailed
examination of factors influencing infant health outcomes.

The data, sourced from birth and death certificates, undergoes a
thorough cleaning and validation process to address inconsistencies and
missing information, ensuring accuracy. Infant mortality rates are
calculated per 1,000 live births, with additional indicators such as
reliability labels for rates based on fewer than 20 deaths. To further
enhance data accuracy, weights are applied to adjust for unlinked infant
death records, ensuring that all deaths—linked and unlinked—are
proportionally represented. Weights are calculated for each state and
infant age-at-death cohort (\<7 days, 7-27 days, and 28 days to 1 year)
using the formula:
$\text{Weight} = \frac{\text{Linked deaths} + \text{Unlinked deaths}}{\text{Linked deaths}}$
This adjustment means areas with complete linkages have a weight of 1.0,
while areas with incomplete linkages are adjusted proportionally.
Additionally, imputation techniques are employed to handle missing
birthweight data, assigning values based on gestational period, race,
sex, and plurality, further reducing potential bias in the analysis.

Once cleaned, the data is stratified by maternal and infant attributes,
providing insights into trends and disparities across demographic
groups. Statistical models are employed to estimate associations between
state-level abortion policies and infant mortality, controlling for
confounders such as maternal characteristics. This structured approach
transforms complex datasets into actionable insights, shedding light on
the health implications of varying abortion laws across the U.S. By
leveraging comprehensive datasets and robust statistical techniques,
this study offers a nuanced understanding of the intersection between
policy, demographics, and infant health outcomes, informing public
health strategies and legislative decisions.

## Outcome variables

### Death Rate

The `death rate` refers to the number of deaths of infants under one
year of age per 1,000 live births in a specified population during a
given time period. It serves as a key indicator of a population’s
overall health and access to healthcare. Rates labeled as "Unreliable"
occur when there are fewer than 20 deaths in the numerator, as these
figures do not meet the National Center for Health Statistics (NCHS)
standards for reliability or precision. After cleaning the dataset, the
"Unreliable" label is converted into a separate variable named
`reliable`.

```{r}
#| label: fig-deathrate
#| fig-cap: The distribution of infant death rate (per 1,000 live births), with a peak observed between 5 and 10 deaths per 1,000 infants.The distribution is right-skewed, highlighting the predominance of lower infant death rates in the dataset.
#| echo: false

ggplot(analysis_data, aes(x = death_rate)) +
  geom_histogram(binwidth = 1, fill = "brown", color = "black") +
  labs(
    x = "Infant Death Rate (per 1,000 live births)", y = "Count",
    caption = "Data source: CDC Wonder."
  ) +theme_classic()
```

@fig-deathrate illustrates the distribution of infant death rates (per
1,000 live births) based on the available data. The majority of infant
death rates cluster around lower values, with a peak observed between 5
and 10 deaths per 1,000 infants. As the death rate increases, the
frequency of occurrences decreases, indicating that higher infant death
rates are relatively rare. There are few instances where the rate
exceeds 20 deaths per 1,000 infants, suggesting these are outliers.
Overall, the distribution is right-skewed, highlighting the predominance
of lower infant death rates in the dataset.

## Predictor variables

### Interaction Term: Abortion Legality and After Injunction

Abortion legality (named `abortion_illegal` in the dataset) is a newly
created variable in our dataset, which refers to the legal status of
abortion within a given state. It is represented as a binary variable,
where 1 indicates that abortion is illegal in the state, and 0 indicates
that abortion is legal. This classification reflects the policies and
laws governing abortion access in each state.

Another created variable `after_injunction` indicates whether the date
of infants' death falls after the implementation of the abortion
injunctions following the overturn of Roe v. Wade in June 2022. It is
coded as a binary variable, where 1 represents dates after June 2022,
signifying the post-injunction period, and 0 represents dates before
June 2022, indicating the pre-injunction period.

To evaluate the impact of abortion injunctions on infant death rates, we
utilize the Difference-in-Difference (DID) method. As part of this
approach, we construct an interaction term defined as the product of
`abortion_illegal` and `after_injunction`. The resulting interaction
term equals 1 when abortion is illegal and the time is after the
injunction, and 0 otherwise. This term is included in the analysis to
evaluate the joint impact of these factors on infant death rates.

```{r}
#| include: false
#| warning: false
#| message: false

library(usmap)
library(ggplot2)
library(dplyr)

# Ensure the number of states and `abortion_illegal` values match
illegal_states <- tolower(c("Alabama", "Arkansas", "Idaho", "Indiana", "Kentucky", 
                    "Louisiana", "Mississippi", "Oklahoma", "South Dakota", 
                    "Tennessee", "Texas", "West Virginia"))
abortion_status <- data.frame(
  state = tolower(c(
    "Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado", "Connecticut", 
    "Delaware", "Florida", "Georgia", "Hawaii", "Idaho", "Illinois", "Indiana", 
    "Iowa", "Kansas", "Kentucky", "Louisiana", "Maine", "Maryland", "Massachusetts", 
    "Michigan", "Minnesota", "Mississippi", "Missouri", "Montana", "Nebraska", 
    "Nevada", "New Hampshire", "New Jersey", "New Mexico", "New York", "North Carolina", 
    "North Dakota", "Ohio", "Oklahoma", "Oregon", "Pennsylvania", "Rhode Island", 
    "South Carolina", "South Dakota", "Tennessee", "Texas", "Utah", "Vermont", 
    "Virginia", "Washington", "West Virginia", "Wisconsin", "Wyoming"
  )),
  short = c(
    "AL", "AK", "AZ", "AR", "CA", "CO", "CT", "DE", "FL", "GA", "HI", "ID", "IL", "IN", 
    "IA", "KS", "KY", "LA", "ME", "MD", "MA", "MI", "MN", "MS", "MO", "MT", "NE", 
    "NV", "NH", "NJ", "NM", "NY", "NC", "ND", "OH", "OK", "OR", "PA", "RI", "SC", 
    "SD", "TN", "TX", "UT", "VT", "VA", "WA", "WV", "WI", "WY"
  )
)|>mutate(abortion_illegal = as.factor(case_when(state %in% illegal_states ~ 1, TRUE ~ 0)))
```

```{r}
#| include: false
#| warning: false
#| message: false
us_map <- map_data("state")
map_data <- us_map|>
   left_join(abortion_status, c("region"="state"))|>filter(!is.na(abortion_illegal))
```

```{r}
#| label: fig-map
#| fig-cap: Abortion Legality by State. As of November 2024, abortion is illegal in Alabama, Arkansas, Idaho, Indiana, Kentucky, Louisiana, Mississippi, Missouri, North Dakota, Oklahoma, South Dakota, Tennessee, Texas, West Virginia, and Wisconsin.
#| echo: false
ggplot() +
  geom_polygon(data = map_data, aes(fill = abortion_illegal, x = long, y = lat, group = group),
    color = "black") +
  scale_fill_manual(
    values = c("0" = "blue", "1" = "red"),
    labels = c("Legal", "Illegal"),
    name = "Abortion Legal Or Not"
  )+
  theme_void() +
  labs(
    caption = "Source: Center for Reproductive Rights\nDate as of Nov.19 2024"
  ) +
  theme(legend.position = "bottom")
```

@fig-map illustrates the legality of abortion across U.S. states. Based
on @reproductiveright, as of November 2024, abortion has been illegal in
Alabama, Arkansas, Idaho, Indiana, Kentucky, Louisiana, Mississippi,
Missouri, North Dakota, Oklahoma, South Dakota, Tennessee, Texas, West
Virginia, and Wisconsin following the June 2022 decision.

```{r}
#| label: fig-rateandstate
#| fig-cap: Infant death rates (per 1,000 live births) over time, with a slight increase after June 2022 in states where abortion is illegal, while rates remain stable in states where abortion is legal.
#| echo: false
#| message: false
analysis_data|>
  mutate(illegal = if_else( abortion_illegal== 1, "Illegal State", "Legal State"))|>
  ggplot(aes(x = date, y = death_rate,colour = illegal)) +
  geom_point(alpha = 0.5)+
  geom_smooth() +
  labs(x = "Date of Infant Death", y = "Infant Death Rate (per 1,000 live births)",color="State is Illegal for Abortion Or Not",caption = "Data source: CDC Wonder.\nIn June 2022, the U.S. Supreme Court overturned Roe v. Wade.") +
  facet_wrap(vars(illegal), nrow = 2)+
  geom_vline(xintercept = as.Date("2022-06-01"), linetype = "dashed")  +
  scale_x_date(
    limits = c(as.Date("2021-01-01"), as.Date("2022-12-01")), 
    date_breaks = "3 months",  
    date_labels = "%Y-%m" 
  )+
  scale_color_brewer(palette = "Set1")+
  theme_classic() +
  theme(legend.position = "bottom")
```

@fig-rateandstate illustrates trends in infant death rates (per 1,000
live births) over time in states where abortion is illegal versus legal.
In states with abortion bans, infant mortality shows a slight increase
after June 2022, marked by the dashed line representing the overturn of
Roe v. Wade. In contrast, states where abortion remains legal exhibit
relatively stable trends. This suggests a potential link between
abortion bans and rising infant mortality rates, warranting further
investigation.

```{r}
#| label: tbl-interact
#| tbl-cap: Average Infant Death Rate (per 1,000 live births) by Interaction Between Abortion Legality and Injunction Timing. States with "Illegal Abortion and After Injunction" (denoted as '1') have a higher average infant death rate compared to states with "Legal Abortion and Before/After Injunction or Illegal Abortion and Before Injunction" (denoted as '0').
#| echo: false
#| warning: false
#| message: false
library(kableExtra)
analysis_data_state <- analysis_data|>mutate(illegal_after_injunction=as.numeric(as.character(abortion_illegal))*as.numeric(as.character(after_injunction)))|>
  group_by(illegal_after_injunction)|>
  summarize(`Average Infant Death Rate (per 1,000 live births)` = round(mean(death_rate, na.rm = TRUE), 1))

kable(analysis_data_state, align = "c") |> 
  footnote(
    "Data Source: CDC Wonder"
  )
```

@tbl-interact highlights average infant death rates (per 1,000 live
births) based on abortion legality and injunction timing. States
categorized as ‘1’, where abortion became illegal after the injunction,
show a higher average infant death rate of 8.3 compared to 7.9 in ‘0’
states, where abortion remained legal or abortion used to be legal
before the injunction. This suggests that restrictive abortion policies
and their enforcement may contribute to higher infant mortality,
potentially due to increased unintended or high-risk pregnancies,
reduced access to maternal healthcare, and exacerbated social or
economic disparities.

## Control Variables

### Year of Death

The Year of Death variable represents the calendar year in which the
infant death occurred, specifically focusing on the years 2021 and 2022
in this study. This variable is essential for analyzing temporal trends
and identifying patterns or shifts in infant mortality rates over this
critical period. By controlling the year of death, we can closely
examine the impact of significant policy changes, such as the
overturning of Roe v. Wade in 2022, as well as other temporal factors on
infant health outcomes.

### State

The state of the mother’s legal residence at the time of birth. This
data is categorized by the mother’s state of residence and excludes
territories.

```{r}
#| label: fig-choromap
#| fig-cap: Infant Death Rates Across U.S. States (per 1,000 Live Births), December 2022. Infant Mortality Rate is much higher iin Southeastern. Alabama and Mississippi show the highest rates, while California and New York exhibit lower rates.
#| echo: false

library(ggnewscale)
death<-analysis_data|>group_by(state)|>summarise(death_rate=mean(death_rate))|>mutate(state=tolower(state))
death_full <- us_map|>
   full_join(death, c("region"="state"))

ggplot() +
  # First layer: Fill with death_rate
  geom_polygon(
    data = death_full,
    aes(x = long, y = lat, group = group, fill = death_rate),
    color = "black"
  ) +
  scale_fill_viridis_c(
    name = "Infant Death Rate\n(per 1,000 live births)",
    option = "C",
    na.value = "grey50",
    breaks = c(10, 15, 20,25),  # Custom breaks for the color legend
  ) +
  theme_void() +
  labs(
    caption = "Grey indicates missing data\nSource: Center for Reproductive Rights\nDate as of December, 2022"
  ) +
  theme(legend.position = "bottom")
```

@fig-choromap is a choropleth map visualizes the average infant death
rate per 1,000 live births across U.S. states, with data categorized by
color intensity. States with higher infant death rates are shown in
lighter shades (yellow), while states with lower rates are represented
in darker shades (purple). States with missing data are indicated in
grey. The map highlights disparities in infant mortality rates, with
some states, particularly in the southeastern U.S., exhibiting notably
higher rates compared to others. States like Alabama and Mississippi
show the highest rates, while states such as California and New York
exhibit lower rates.

### Age of Mother

The Age of Mother (`age_of_mother`) refers to the age group of the
mother at the time of childbirth.In our observational data, age is
categorized into the brackets 15–19 years, 20–24 years, 25–29 years, and
30–34 years in the raw data, in compliance with privacy requirements.

@tbl-motherage displays the average infant death rate (per 1,000 live
births) by the age group of mothers. It shows a clear trend: younger
mothers, particularly those aged 15–19 and 20–24, have higher infant
death rates (9.0 and 10.0, respectively), while older mothers in the
25–29 and 30–34 age groups have lower rates (7.4 and 6.6, respectively).
This suggests that maternal age may be a factor influencing infant
mortality, with younger mothers experiencing higher risks.

```{r}
#| label: tbl-motherage
#| tbl-cap: Average Infant Death Rate (per 1,000 live births) of different age group of Mother. It shows that infant death rates decrease with increasing maternal age, with the highest rate (10.0) among mothers aged 20–24 and the lowest (6.6) among those aged 30–34.
#| echo: false
#| warning: false
#| message: false
library(kableExtra)
analysis_data_state <- analysis_data|>
  group_by(age_of_mother)|>
  summarize(`Average Infant Death Rate (per 1,000 live births)` = round(mean(death_rate, na.rm = TRUE), 1)) |>rename(`Age of Mother`=age_of_mother)

# Display the result as a table
kable(analysis_data_state)|>footnote("Data Source: CDC Wonder")
```

### Mother's Single Race

The self-identified race of the mother, recorded as a single race
category. In this dataset, the available categories are: Asian, Black or
African American, and White. This classification facilitates the
analysis of maternal and infant health outcomes across these racial
groups.

```{r}
#| label: tbl-motherrace
#| tbl-cap: Average Infant Death Rate (per 1,000 live births) of different age group of Mother. It shows that infant death rates decrease with increasing maternal age, with the highest rate (10.0) among mothers aged 20–24 and the lowest (6.6) among those aged 30–34.
#| echo: false
#| warning: false
#| message: false
analysis_data_state <- analysis_data|>
  group_by(mothers_single_race)|>
  summarize(`Average Infant Death Rate (per 1,000 live births)` = round(mean(death_rate, na.rm = TRUE), 1)) |>rename(`Mother's Single Race`=mothers_single_race)

# Display the result as a table
kable(analysis_data_state)|>footnote("Data Source: CDC Wonder")
```

# Model

For our analysis, we use a Bayesian Difference-in-Differences (DID) model to quantify the impact of abortion injunctions on the infant death rate across states in the USA. This method compares changes in infant death rates over time between a treatment group (states where abortion injunctions were implemented) and a control group (states where abortion remains legal), providing a robust framework for estimating causal effects.

The model focuses on the infant death rate per 1,000 live births as the dependent variable and includes the following independent variables:\
	-	Treatment Group Indicator `abortion_illegal`: A binary variable indicating the group: 1 if abortion is illegal in the state. 0 if abortion remains legal.\
	-	Time Indicator `after_injunction`: A binary variable indicating the time period: 0 for the period before the injunction implemented. 1 for the period after the injunction implemented.
	-	Interaction Term`after_injunction` × `abortion_illegal`: The interaction term captures the additional change in the outcome (e.g., percentage of support, behavior, or health-related metrics) for regions where abortion is illegal after the injunction implemented, beyond any changes observed in regions where abortion remains legal.

To account for potential regional variations in infant death rates, we include state as a random effect in the model. This allows us to control for unobserved heterogeneity across states that might introduce bias. Additionally, we include age_of_mother, mothers_single_race, and year_of_death as control variables to adjust for demographic and temporal factors that could influence infant mortality rates, ensuring that the estimated effects are not confounded by these characteristics.

By isolating the impact of the abortion injunction through this interaction term, the model identifies changes in infant death rates that can be attributed to the policy intervention. Additional details and diagnostics supporting the model’s implementation are provided in [Appendix-@sec-model-details].

## Difference-In-Difference Approach (DID)

DID is a causal inference method that estimates the impact of an intervention by comparing changes in outcomes over time between a treatment group and a control group. By assuming parallel trends in the absence of the intervention, the DID framework isolates the intervention’s effect through the interaction between time periods (pre- and post-intervention) and group membership. This makes it a robust method for evaluating policy changes and other significant events.

## Model set-up

In our analysis, we employ two models. The first model includes only the predictor variables and the random effect for state. The second model builds on the first by incorporating additional control variables to account for potential confounding factors.

Define $y_i$ as as the infant death rate per 1,000 live births.
$$y_i|\mu_i,\sigma \sim \mbox{Normal}(\mu_i, \sigma)\\
\text{First Model}: \mu_i = \beta_0 + \beta_1\cdot\text{After Injunction} + \beta_2\cdot\text{Abortion Illegal}+ \beta_3\cdot\text{After Injunction}\cdot\text{Abortion Illegal}_i+\gamma_j\\
\text{Second Model}:\mu_i = \beta_0 + \beta_1\cdot\text{After Injunction}_i + \beta_2\cdot\text{Abortion Illegal}_i+ \beta_3\cdot\text{After Injunction}_i\cdot\text{Abortion Illegal}_i\\+\beta_4\cdot\text{Year of Death}_i+\beta_5\cdot\text{Age of Mother}_i+\beta_6\cdot\text{Mother's Single Race}_i+\gamma_j\\
\beta_0 \sim \mbox{Normal}(0, 2.5) \\
\beta_1 \sim \mbox{Normal}(0, 2.5) \\
\beta_2 \sim \mbox{Normal}(0, 2.5) \\
\beta_3 \sim \mbox{Normal}(0, 2.5) \\
\beta_4 \sim \mbox{Normal}(0, 2.5) \\
\beta_5 \sim \mbox{Normal}(0, 2.5) \\
\beta_6 \sim \mbox{Normal}(0, 2.5) \\
\gamma_j \sim \text{Normal}(0, \sigma_j^2)\\
\sigma \sim \mbox{Exponential}(1)$$

The models are run in R (R Core Team 2023) using the rstanarm package of @rstanarm. The default priors from `rstanarm` are used for both GLM Bayesian models. We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm` used for both Gaussian Bayesian models.

### Model justification

The Bayesian framework enhances the DID approach by incorporating prior information and quantifying uncertainty in a probabilistic manner. Using Markov Chain Monte Carlo (MCMC) sampling, the model estimates the posterior distributions of the parameters, offering flexibility in handling complex relationships and uncertainties in the data.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| include: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
second_model <-
  readRDS(file = here::here("models/second_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::msummary(
  list(first_model,second_model),
  fmt = 3,coef_omit="state",notes = "Note. All models include a control for state and pollsters. The reference level of the state is Alaska, the reference level of population is likely voter, and the reference level of pollsters is Angus Reid.")
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The
discussion is a chance to show off what you know and what you learnt
from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what
your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {.unnumbered}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive
check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the
prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

```{r}
plot(first_model, pars = "(Intercept)", prob = 0.95)
```

```{r}
prior_summary(second_model)
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

```{r}
plot(
  first_model,
  "areas"
)
plot(
  second_model,
  "areas"
)
```

```{r}
pp_check(second_model) +
  theme_classic() +
  theme(legend.position = "bottom")
```

\newpage

# References