### 2024-11-24 paper/paper.qmd
update paper

---
title: "Examining the impact Overturn of Roe v. Wade: Banning of Abortion on Infant Mortality Rates in the United States Using a Difference-in-Differences Approach"
subtitle: "The infant death rate increased by 0.285 deaths per 1,000 live births in states that became abortion-banned, compared to states where abortion remained legal."
author: 
  - Diana Shen
thanks: "Code and data are available at: [https://github.com/DianaShen1224/Relationship-between-infant-mortality-rate-and-prohibited-abortion)."
date: today
date-format: long
abstract: "This study examines the impact of the Dobbs decision and subsequent abortion bans on infant mortality rates in the United States, focusing on data from 2021 to 2022. Using a Difference-in-Differences (DID) approach, we analyze trends in states where abortion remained legal compared to those where it became illegal after June 2022. Our results indicate that states enforcing abortion bans experienced an increase of 0.285 infant deaths per 1,000 live births, with compounded effects observed during the post-injunction period. Maternal age and race also emerged as significant predictors, with Black mothers facing disproportionately higher infant mortality rates. These findings underscore the compounded public health implications of restrictive abortion policies and the urgent need for targeted interventions to address racial and demographic disparities. By utilizing recent, high-quality data and rigorous modeling, this study provides critical insights into the intersection of reproductive rights and infant health outcomes."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
editor: 
  markdown: 
    wrap: 72
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(arrow)
library(knitr)
```

```{r}
#| include: false
#| warning: false
#| message: false
analysis_data<-read_parquet(here::here("data", "02-analysis_data", "analysis_data.parquet"))
```

# Introduction

The overturning of Roe v. Wade on June 24, 2022, through the Supreme
Court’s decision in Dobbs v. Jackson Women’s Health Organization, ended
nearly five decades of constitutional protection for abortion rights.
This landmark ruling shifted authority over abortion regulations to
individual states, creating a patchwork of laws nationwide. The decision
reignited debates about reproductive rights and public health,
highlighting disparities in access to abortion services and their
effects on health outcomes, especially in states with restrictive
policies.

Banning abortion usually has been linked to a rise in maternal mortality
rates, with Texas experiencing a 56% increase from 2019 to 2022
[@nbcnews]. However, restrictive abortion laws can also impact infant
mortality, as they often lead to more unintended or high-risk
pregnancies among disadvantaged populations with limited access to
healthcare. These pregnancies are associated with higher rates of
preterm births, low birth weights, and delivery complications, all of
which elevate infant mortality rates. Even before Roe v. Wade was
overturned, researchers highlighted the impacts of restrictive abortion
laws on mortality rates. @Burdick2024 found that states with 11–12
restrictive laws during 2014–2018 had a 16% higher infant mortality
rate, particularly in counties with higher proportions of Black
populations, inadequate prenatal care, and maternal smoking. Similarly,
@Harper2023 showed that restrictive laws are linked to higher maternal,
fetal, and infant mortality, states with moderate abortion legislation
had significantly lower maternal mortality (25.79 per 100,000 live
births) and infant mortality (20.56 per 1,000 live births) compared to
restrictive states, underscoring the health benefits of less restrictive
policies.

The estimand of our analysis is the change in the difference in infant
death rates between states that banned abortion and states where
abortion remained legal after June 24, 2022. The object of the
estimation is to estimate the change in the difference in the infant
death rate between the treatment and control groups based on
observational data in the United States.

Instead of examining the impact of abortion laws over broader time
periods like prior research, our study focuses on the effect of abortion
bans on infant mortality rates in the United States specifically from
2021 to 2022, with an emphasis on changes following the 2022 Dobbs
decision. Using a Difference-in-Differences (DID) approach, we assess
the interaction between abortion legality in states and the timing of
the injunction implemented to understand its influence on infant health
outcomes. The study utilizes observational data from the Centers for
Disease Control and Prevention’s WONDER database, capturing infant
mortality trends across states with varying levels of abortion
restrictiveness. Key predictors include the mother’s state of residence,
race, age, and the year of infant death, enabling a nuanced examination
of the effects of restrictive abortion laws on public health.

Our Results show that the overturning of Roe v. Wade and the
implementation of abortion bans are linked to a rise in infant mortality
rates in states where abortion became illegal. Through a
Difference-in-Differences (DID) approach, we estimate that states
enforcing abortion bans after the injunction saw an increase of 0.285
deaths per 1,000 live births compared to states where abortion remained
legal. Furthermore, we find that maternal age and race are key
determinants of infant mortality, with older maternal age groups
associated with lower mortality rates and Black mothers experiencing
significantly higher rates. These results emphasize the public health
consequences of restrictive abortion laws and the urgent need for
policies that mitigate these disparities.

This study is important as we highlight the public health implications
of abortion bans, particularly their role in increasing infant mortality
rates. By focusing on the 2022 Dobbs decision, we provide timely
evidence on how restrictive abortion policies disproportionately affect
vulnerable populations, emphasizing the need for data-driven policies to
mitigate these adverse outcomes.

The structure of this paper is organized as follows: @sec-data provides
details on the data sources and variables used in our analysis.
@sec-model explains the modeling approach, including the assumptions and
specifications of our linear regression framework. In @sec-result, we
present our findings, emphasizing the key predictors of Harris's
support. Finally, @sec-discussion explores the implications of our
results and suggests potential directions for future research.
[Appendix- @sec-data-details] provides external data detail, [Appendix-
@sec-model-details] provides model detail, @sec-c provides an datasheet,
and @sec-d provides an detailed introduction of surveys, sampling and
observational data in our paper.

# Data {#sec-data}

## Overview

We conduct our infant death rate observational data analysis using the R
programming language [@citeR]. Our dataset, obtained from CDC
Wonder[@cdc_wonder_2024], based on data as of December 2022, provides a
detailed overview of the Infant Death Rate across the USA. CDC WONDER
(Wide-ranging ONline Data for Epidemiologic Research) is an online
system providing public health professionals and the public with access
to a wide range of the Centers for Disease Control and Prevention (CDC)
epidemiologic data and resources. Following the guidelines outlined in
@tellingstories, we analyze the impact of abortion injunctions on infant
death rates using the Difference-in-Difference method. This approach
utilizes the interaction between the legality of abortion in the
mother’s state of residence and the timing of the infant’s death as the
predictor variable. Additionally, we include control variables such as
infant year of death, the mother’s state of residence, race, and age to
account for potential confounding factors.

## Measurement

The process of translating real-world events into our dataset requires a
systematic approach to measurement and data gathering. In this research,
we analyze infant mortality rates in the United States to understand the
impact of abortion policies and demographic factors. The CDC’s Linked
Birth/Infant Death Records database provides comprehensive data,
including maternal and infant characteristics, to support this study.
Variables such as maternal age, race, education, birth weight,
gestational age, and prenatal care are included to ensure a detailed
examination of factors influencing infant health outcomes.

The data, sourced from birth and death certificates, undergoes a
thorough cleaning and validation process to address inconsistencies and
missing information, ensuring accuracy. Infant mortality rates are
calculated per 1,000 live births, with additional indicators such as
reliability labels for rates based on fewer than 20 deaths. To further
enhance data accuracy, weights are applied to adjust for unlinked infant
death records, ensuring that all deaths—linked and unlinked—are
proportionally represented. Weights are calculated for each state and
infant age-at-death cohort (\<7 days, 7-27 days, and 28 days to 1 year)
using the formula:
$\text{Weight} = \frac{\text{Linked deaths} + \text{Unlinked deaths}}{\text{Linked deaths}}$
This adjustment means areas with complete linkages have a weight of 1.0,
while areas with incomplete linkages are adjusted proportionally.
Additionally, imputation techniques are employed to handle missing
birthweight data, assigning values based on gestational period, race,
sex, and plurality, further reducing potential bias in the analysis.

Once cleaned, the data is stratified by maternal and infant attributes,
providing insights into trends and disparities across demographic
groups. Statistical models are employed to estimate associations between
state-level abortion policies and infant mortality, controlling for
confounders such as maternal characteristics. This structured approach
transforms complex datasets into actionable insights, shedding light on
the health implications of varying abortion laws across the U.S. By
leveraging comprehensive datasets and robust statistical techniques,
this study offers a nuanced understanding of the intersection between
policy, demographics, and infant health outcomes, informing public
health strategies and legislative decisions.

## Outcome variables

### Death Rate

The `death rate` refers to the number of deaths of infants under one
year of age per 1,000 live births in a specified population during a
given time period. It serves as a key indicator of a population’s
overall health and access to healthcare. Rates labeled as "Unreliable"
occur when there are fewer than 20 deaths in the numerator, as these
figures do not meet the National Center for Health Statistics (NCHS)
standards for reliability or precision. After cleaning the dataset, the
"Unreliable" label is converted into a separate variable named
`reliable`.

```{r}
#| label: fig-deathrate
#| fig-cap: The distribution of infant death rate (per 1,000 live births), with a peak observed between 5 and 10 deaths per 1,000 infants.The distribution is right-skewed, highlighting the predominance of lower infant death rates in the dataset.
#| echo: false

ggplot(analysis_data, aes(x = death_rate)) +
  geom_histogram(binwidth = 1, fill = "brown", color = "black") +
  labs(
    x = "Infant Death Rate (per 1,000 live births)", y = "Count",
    caption = "Data source: CDC Wonder."
  ) +theme_classic()
```

@fig-deathrate illustrates the distribution of infant death rates (per
1,000 live births) based on the available data. The majority of infant
death rates cluster around lower values, with a peak observed between 5
and 10 deaths per 1,000 infants. As the death rate increases, the
frequency of occurrences decreases, indicating that higher infant death
rates are relatively rare. There are few instances where the rate
exceeds 20 deaths per 1,000 infants, suggesting these are outliers.
Overall, the distribution is right-skewed, highlighting the predominance
of lower infant death rates in the dataset.

## Predictor variables

### Interaction Term: Abortion Legality and After Injunction

Abortion legality (named `abortion_illegal` in the dataset) is a newly
created variable in our dataset, which refers to the legal status of
abortion within a given state. It is represented as a binary variable,
where 1 indicates that abortion is illegal in the state, and 0 indicates
that abortion is legal. This classification reflects the policies and
laws governing abortion access in each state.

Another created variable `after_injunction` indicates whether the date
of infants' death falls after the implementation of the abortion
injunctions following the overturn of Roe v. Wade in June 2022. It is
coded as a binary variable, where 1 represents dates after June 2022,
signifying the post-injunction period, and 0 represents dates before
June 2022, indicating the pre-injunction period.

To evaluate the impact of abortion injunctions on infant death rates, we
utilize the Difference-in-Difference (DID) method. As part of this
approach, we construct an interaction term defined as the product of
`abortion_illegal` and `after_injunction`. The resulting interaction
term equals 1 when abortion is illegal and the time is after the
injunction, and 0 otherwise. This term is included in the analysis to
evaluate the joint impact of these factors on infant death rates.

```{r}
#| include: false
#| warning: false
#| message: false

library(usmap)
library(ggplot2)
library(dplyr)

# Ensure the number of states and `abortion_illegal` values match
illegal_states <- tolower(c("Alabama", "Arkansas", "Idaho", "Indiana", "Kentucky", 
                    "Louisiana", "Mississippi", "Oklahoma", "South Dakota", 
                    "Tennessee", "Texas", "West Virginia"))
abortion_status <- data.frame(
  state = tolower(c(
    "Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado", "Connecticut", 
    "Delaware", "Florida", "Georgia", "Hawaii", "Idaho", "Illinois", "Indiana", 
    "Iowa", "Kansas", "Kentucky", "Louisiana", "Maine", "Maryland", "Massachusetts", 
    "Michigan", "Minnesota", "Mississippi", "Missouri", "Montana", "Nebraska", 
    "Nevada", "New Hampshire", "New Jersey", "New Mexico", "New York", "North Carolina", 
    "North Dakota", "Ohio", "Oklahoma", "Oregon", "Pennsylvania", "Rhode Island", 
    "South Carolina", "South Dakota", "Tennessee", "Texas", "Utah", "Vermont", 
    "Virginia", "Washington", "West Virginia", "Wisconsin", "Wyoming"
  )),
  short = c(
    "AL", "AK", "AZ", "AR", "CA", "CO", "CT", "DE", "FL", "GA", "HI", "ID", "IL", "IN", 
    "IA", "KS", "KY", "LA", "ME", "MD", "MA", "MI", "MN", "MS", "MO", "MT", "NE", 
    "NV", "NH", "NJ", "NM", "NY", "NC", "ND", "OH", "OK", "OR", "PA", "RI", "SC", 
    "SD", "TN", "TX", "UT", "VT", "VA", "WA", "WV", "WI", "WY"
  )
)|>mutate(abortion_illegal = as.factor(case_when(state %in% illegal_states ~ 1, TRUE ~ 0)))
```

```{r}
#| include: false
#| warning: false
#| message: false
us_map <- map_data("state")
map_data <- us_map|>
   left_join(abortion_status, c("region"="state"))|>filter(!is.na(abortion_illegal))
```

```{r}
#| label: fig-map
#| fig-cap: Abortion Legality by State. As of November 2024, abortion is illegal in Alabama, Arkansas, Idaho, Indiana, Kentucky, Louisiana, Mississippi, Missouri, North Dakota, Oklahoma, South Dakota, Tennessee, Texas, West Virginia, and Wisconsin.
#| echo: false
ggplot() +
  geom_polygon(data = map_data, aes(fill = abortion_illegal, x = long, y = lat, group = group),
    color = "black") +
  scale_fill_manual(
    values = c("0" = "blue", "1" = "red"),
    labels = c("Legal", "Illegal"),
    name = "Abortion Legal Or Not"
  )+
  theme_void() +
  labs(
    caption = "Source: Center for Reproductive Rights\nDate as of Nov.19 2024"
  ) +
  theme(legend.position = "bottom")
```

@fig-map illustrates the legality of abortion across U.S. states. Based
on @reproductiveright, as of November 2024, abortion has been illegal in
Alabama, Arkansas, Idaho, Indiana, Kentucky, Louisiana, Mississippi,
Missouri, North Dakota, Oklahoma, South Dakota, Tennessee, Texas, West
Virginia, and Wisconsin following the June 2022 decision.

```{r}
#| label: fig-rateandstate
#| fig-cap: Infant death rates (per 1,000 live births) over time, with a slight increase after June 2022 in states where abortion is illegal, while rates remain stable in states where abortion is legal.
#| echo: false
#| message: false
analysis_data|>
  mutate(illegal = if_else( abortion_illegal== 1, "Illegal State", "Legal State"))|>
  ggplot(aes(x = date, y = death_rate,colour = illegal)) +
  geom_point(alpha = 0.5)+
  geom_smooth() +
  labs(x = "Date of Infant Death", y = "Infant Death Rate (per 1,000 live births)",color="State is Illegal for Abortion Or Not",caption = "Data source: CDC Wonder.\nIn June 2022, the U.S. Supreme Court overturned Roe v. Wade.") +
  facet_wrap(vars(illegal), nrow = 2)+
  geom_vline(xintercept = as.Date("2022-06-01"), linetype = "dashed")  +
  scale_x_date(
    limits = c(as.Date("2021-01-01"), as.Date("2022-12-01")), 
    date_breaks = "3 months",  
    date_labels = "%Y-%m" 
  )+
  scale_color_brewer(palette = "Set1")+
  theme_classic() +
  theme(legend.position = "bottom")
```

@fig-rateandstate illustrates trends in infant death rates (per 1,000
live births) over time in states where abortion is illegal versus legal.
In states with abortion bans, infant mortality shows a slight increase
after June 2022, marked by the dashed line representing the overturn of
Roe v. Wade. In contrast, states where abortion remains legal exhibit
relatively stable trends. This suggests a potential link between
abortion bans and rising infant mortality rates, warranting further
investigation.

```{r}
#| label: tbl-interact
#| tbl-cap: Average Infant Death Rate (per 1,000 live births) by Interaction Between Abortion Legality and Injunction Timing. States with "Illegal Abortion and After Injunction" (denoted as '1') have a higher average infant death rate compared to states with "Legal Abortion and Before/After Injunction or Illegal Abortion and Before Injunction" (denoted as '0').
#| echo: false
#| warning: false
#| message: false
library(kableExtra)
analysis_data_state <- analysis_data|>mutate(illegal_after_injunction=as.numeric(as.character(abortion_illegal))*as.numeric(as.character(after_injunction)))|>
  group_by(illegal_after_injunction)|>
  summarize(`Average Infant Death Rate (per 1,000 live births)` = round(mean(death_rate, na.rm = TRUE), 1))

kable(analysis_data_state, align = "c") |> 
  footnote(
    "Data Source: CDC Wonder"
  )
```

@tbl-interact highlights average infant death rates (per 1,000 live
births) based on abortion legality and injunction timing. States
categorized as ‘1’, where abortion became illegal after the injunction,
show a higher average infant death rate of 8.3 compared to 7.9 in ‘0’
states, where abortion remained legal or abortion used to be legal
before the injunction. This suggests that restrictive abortion policies
and their enforcement may contribute to higher infant mortality,
potentially due to increased unintended or high-risk pregnancies,
reduced access to maternal healthcare, and exacerbated social or
economic disparities.

## Control Variables

### Year of Death

The Year of Death variable represents the calendar year in which the
infant death occurred, specifically focusing on the years 2021 and 2022
in this study. This variable is essential for analyzing temporal trends
and identifying patterns or shifts in infant mortality rates over this
critical period. By controlling the year of death, we can closely
examine the impact of significant policy changes, such as the
overturning of Roe v. Wade in 2022, as well as other temporal factors on
infant health outcomes.

### Age of Mother

The Age of Mother (`age_of_mother`) refers to the age group of the
mother at the time of childbirth.In our observational data, age is
categorized into the brackets 15–19 years, 20–24 years, 25–29 years, and
30–34 years in the raw data, in compliance with privacy requirements.

@tbl-motherage displays the average infant death rate (per 1,000 live
births) by the age group of mothers. It shows a clear trend: younger
mothers, particularly those aged 15–19 and 20–24, have higher infant
death rates (9.0 and 10.0, respectively), while older mothers in the
25–29 and 30–34 age groups have lower rates (7.4 and 6.6, respectively).
This suggests that maternal age may be a factor influencing infant
mortality, with younger mothers experiencing higher risks.

```{r}
#| label: tbl-motherage
#| tbl-cap: Average Infant Death Rate (per 1,000 live births) of different age group of Mother. It shows that infant death rates decrease with increasing maternal age, with the highest rate (10.0) among mothers aged 20–24 and the lowest (6.6) among those aged 30–34.
#| echo: false
#| warning: false
#| message: false
library(kableExtra)
analysis_data_state <- analysis_data|>
  group_by(age_of_mother)|>
  summarize(`Average Infant Death Rate (per 1,000 live births)` = round(mean(death_rate, na.rm = TRUE), 1)) |>rename(`Age of Mother`=age_of_mother)

# Display the result as a table
kable(analysis_data_state)|>footnote("Data Source: CDC Wonder")
```

### Mother's Single Race

The self-identified race of the mother, recorded as a single race
category. In this dataset, the available categories are: Asian, Black or
African American, and White. This classification facilitates the
analysis of maternal and infant health outcomes across these racial
groups.

```{r}
#| label: tbl-motherrace
#| tbl-cap: Average Infant Death Rate (per 1,000 live births) of different age group of Mother. It shows that infant death rates decrease with increasing maternal age, with the highest rate (10.0) among mothers aged 20–24 and the lowest (6.6) among those aged 30–34.
#| echo: false
#| warning: false
#| message: false
analysis_data_state <- analysis_data|>
  group_by(mothers_single_race)|>
  summarize(`Average Infant Death Rate (per 1,000 live births)` = round(mean(death_rate, na.rm = TRUE), 1)) |>rename(`Mother's Single Race`=mothers_single_race)

# Display the result as a table
kable(analysis_data_state)|>footnote("Data Source: CDC Wonder")
```

## Random Variable

### State

The state of the mother’s legal residence at the time of birth. This
data is categorized by the mother’s state of residence and excludes
territories.

```{r}
#| label: fig-choromap
#| fig-cap: Infant Death Rates Across U.S. States (per 1,000 Live Births), December 2022. Infant Mortality Rate is much higher iin Southeastern. Alabama and Mississippi show the highest rates, while California and New York exhibit lower rates.
#| echo: false

library(ggnewscale)
death<-analysis_data|>group_by(state)|>summarise(death_rate=mean(death_rate))|>mutate(state=tolower(state))
death_full <- us_map|>
   full_join(death, c("region"="state"))

ggplot() +
  # First layer: Fill with death_rate
  geom_polygon(
    data = death_full,
    aes(x = long, y = lat, group = group, fill = death_rate),
    color = "black"
  ) +
  scale_fill_viridis_c(
    name = "Infant Death Rate\n(per 1,000 live births)",
    option = "C",
    na.value = "grey50",
    breaks = c(10, 15, 20,25),  # Custom breaks for the color legend
  ) +
  theme_void() +
  labs(
    caption = "Grey indicates missing data\nSource: Center for Reproductive Rights\nDate as of December, 2022"
  ) +
  theme(legend.position = "bottom")
```

@fig-choromap is a choropleth map visualizes the average infant death
rate per 1,000 live births across U.S. states, with data categorized by
color intensity. States with higher infant death rates are shown in
lighter shades (yellow), while states with lower rates are represented
in darker shades (purple). States with missing data are indicated in
grey. The map highlights disparities in infant mortality rates, with
some states, particularly in the southeastern U.S., exhibiting notably
higher rates compared to others. States like Alabama and Mississippi
show the highest rates, while states such as California and New York
exhibit lower rates. \# Model

For our analysis, we use a Bayesian Difference-in-Differences (DID)
model to quantify the impact of abortion injunctions on the infant death
rate across states in the USA. This method compares changes in infant
death rates over time between a treatment group (states where abortion
injunctions were implemented) and a control group (states where abortion
remains legal), providing a robust framework for estimating causal
effects.

## Difference-In-Difference Approach (DID)

Difference-in-Differences (DID) is a causal inference method that
estimates the impact of an intervention by comparing changes in outcomes
over time between a treatment group and a control group. By assuming
parallel trends in the absence of the intervention, the DID framework
isolates the intervention’s effect through **the interaction between
time periods (pre- and post-intervention) and group membership**. This
makes it a robust method for evaluating policy changes and other
significant events.

### Assumption

The validity of the Difference-in-Differences (DID) approach in our
study relies on several key assumptions, which, if violated, could
affect the robustness and reliability of our results [@cunningham2021]:

#### Parallel Trends Assumption

The model relies on the assumption that, without the abortion
injunction, the treatment group (states where abortion became illegal)
and the control group (states where abortion remained legal) would have
followed similar trends in infant mortality rates. However, this
assumption can be challenged by differences in healthcare systems,
socioeconomic conditions, or pre-existing policy environments, which may
cause deviations from parallel trends.

#### Compositional Differences

In addition to parallel trends, compositional changes in repeated
cross-sectional data over time can influence results. For instance,
shifts in demographic factors such as maternal age or racial makeup
between pre- and post-intervention periods may confound the observed
changes in infant mortality rates. Including control variables helps
adjust for these confounders, ensuring that the estimated effect
reflects the policy intervention rather than changes in population
composition.

#### Long-term Effects vs. Reliability

The length of the analysis period further complicates the reliability of
results. While longer periods allow for capturing long-term effects,
they also increase the likelihood of unrelated factors, such as new
policies or healthcare advancements, influencing the outcome. Over time,
some states in the control group may also implement similar abortion
policies, diluting the treatment effect. Conversely, shorter analysis
periods

#### Functional Form Dependence

Finally, the model assumes that the functional relationship between
predictors and outcomes is correctly specified. If this assumption is
violated, differences in outcomes between treatment and control groups
may stem from misspecified relationships rather than the intervention
itself. This highlights the importance of carefully assessing the
model’s structure and assumptions to ensure valid results.

## Model Parameters

The model focuses on the infant death rate per 1,000 live births as the
dependent variable and includes the following independent variables:

-   **Treatment Group Indicator `abortion_illegal`**: A binary variable
    indicating the group: 1 if abortion is illegal in the state. 0 if
    abortion remains legal.

-   **Time Indicator `after_injunction`**: A binary variable indicating
    the time period: 0 for the period before the injunction implemented.
    1 for the period after the injunction implemented.

-   **Interaction Term`after_injunction` × `abortion_illegal`**: The
    interaction term captures the additional change in the outcome
    (e.g., percentage of support, behavior, or health-related metrics)
    for regions where abortion is illegal after the injunction
    implemented, beyond any changes observed in regions where abortion
    remains legal.

Additionally, we include these control variables to adjust for
demographic and temporal factors that could influence infant mortality
rates, ensuring that the estimated effects are not confounded by these
characteristics:

-   **`age_of_mother`**: The age of the mother at the time of birth,
    represented as a continuous variable.

-   **`mothers_single_race`**: Self-identified race of the mother,
    represented as a categorical variable.

-   **`year_of_death`**: The year in which the infant death occurred,
    represented as a continuous variable.

Finally, we include a random variable:

-   **`state`**: Mother's residence state when the infant born as a
    random effect in the model.

By isolating the impact of the abortion injunction through this
interaction term `after_injunction`×`abortion_illegal`, our model
identifies changes in infant death rates that can be attributed to the
policy intervention. Additional details and diagnostics supporting the
model’s implementation are provided in [Appendix- @sec-model-details].

## Model set-up

In our analysis, we employ two models. The first model includes only the
predictor variables and the random effect for state. The second model
builds on the first by incorporating additional control variables to
account for potential confounding factors.

Define $y_i$ as as the infant death rate per 1,000 live births, and
$\gamma_j$ as the random state effect. $$
\begin{aligned}
y_i|\mu_i,\sigma &\sim \mbox{Normal}(\mu_i, \sigma)\\
\text{First Model}: \mu_i &= \beta_0 + \beta_1\cdot\text{After Injunction} + \beta_2\cdot\text{Abortion Illegal}\\&+\beta_3\cdot\text{After Injunction}\cdot\text{Abortion Illegal}_i+\gamma_j\\
\text{Second Model}:\mu_i &= \beta_0 + \beta_1\cdot\text{After Injunction}_i + \beta_2\cdot\text{Abortion Illegal}_i\\&+ \beta_3\cdot\text{After Injunction}_i\cdot\text{Abortion Illegal}_i\\&+\beta_4\cdot\text{Year of Death}_i+\beta_5\cdot\text{Age of Mother}_i+\beta_6\cdot\text{Mother's Single Race}_i+\gamma_j\\
\beta_0 &\sim \mbox{Normal}(0, 2.5)\\
\beta_1 &\sim \mbox{Normal}(0, 2.5)\\
\beta_2 &\sim \mbox{Normal}(0, 2.5)\\
\beta_3 &\sim \mbox{Normal}(0, 2.5)\\
\beta_4 &\sim \mbox{Normal}(0, 2.5)\\
\beta_5 &\sim \mbox{Normal}(0, 2.5)\\
\beta_6 &\sim \mbox{Normal}(0, 2.5)\\
\gamma_j&\sim \text{Normal}(0, \sigma_j^2)\\
\sigma &\sim \mbox{Exponential}(1)
\end{aligned}$$

The models are run in R (R Core Team 2023) using the rstanarm package of
@rstanarm. The default priors from `rstanarm` are used for both GLM
Bayesian models. We run the model in R [@citeR] using the `rstanarm`
package of @rstanarm. We use the default priors from `rstanarm` used for
both Gaussian Bayesian models.

### Model justification

For our analysis, we use separate Bayesian regression models to estimate
the infant death rates in states where abortion injunctions were
implemented and in states where abortion remains legal. We use the
interaction term between after_injunction and abortion_illegal to
measure the causal effect of the abortion injunction. Both models share
the same predictors to ensure consistency and comparability in capturing
the determinants of infant mortality.

For one of our models, we include control variables (`age_of_mother`,
`mothers_single_race`, and `year_of_death`) to adjust for confounders
that could influence infant death rates, addressing the assumption of
compositional differences in the DID approach and ensuring unbiased
estimates of the intervention’s effect. In the other model, we exclude
these controls to focus on the overall intervention impact, avoid
overfitting, and prioritize simplicity when variability is adequately
accounted for by the design or random effects. Thus, our decision to fit
two models is to compare the outcomes: one adjusts for confounders to
refine the estimates, while the other focuses on the direct effect of
the intervention.

Our model includes `state` as a random effect to account for regional
variations in infant mortality that could introduce bias. We use a
random effect rather than a fixed effect to capture variability across
states without estimating specific effects for each one. This approach
models unobserved heterogeneity, such as differences in healthcare
systems and socioeconomic conditions, using a single variance term,
which avoids overfitting and ensures efficiency and generalizability. By
adjusting for unmeasured state-specific factors, the model provides
robust and reliable findings while balancing parsimony and
interpretability, making it well-suited for our analysis.

We choose the Bayesian framework since it enhances the DID approach by
incorporating prior information and quantifying uncertainty in a
probabilistic manner. In our Bayesian models, we use weakly informative
priors from the `rstanarm` package, including a Normal prior ($\mu=0$
and $\sigma=2.5$) for fixed effects and an Exponential prior (rate = 1)
for the auxiliary parameter, with `autoscale=TRUE` to dynamically adjust
priors based on data scale. This enhances convergence and ensures priors
are calibrated to predictor ranges without dominating the data. The
model assumes plausible parameter centering, random effects capture
unobserved heterogeneity, and the DID framework relies on parallel
trends between groups. However, limitations include potential bias from
misspecified priors, computational challenges for hierarchical models,
and sensitivity to data sparsity or imbalances. This model may not be
suitable if group-level effects are negligible, parallel trends are
violated, or data is insufficient, but when applied appropriately, it
provides robust and interpretable causal estimates.

# Results

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
second_model <-
  readRDS(file = here::here("models/second_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::msummary(
  list(first_model,second_model),
  fmt = 3,coef_omit="state",notes = "Note. All models include a random variable for state. The reference level of age_of_mother is 15-19.The reference level of monthers_single_race is Asian.")
```

The results from the two Bayesian models are summarized in
@tbl-modelresults. The interaction term between after_injunction and
abortion_illegal is positive and statistically significant (0.111 in
Model 1 and 0.280 in Model 2), indicating a change in infant mortality
rates associated with the combination of abortion bans and the
post-injunction period.

For the independent effects, states where abortion is illegal show
higher infant mortality rates, with coefficients of 3.365 in Model 1 and
1.400 in Model 2. The post-injunction period alone is associated with
coefficients of -0.525 in Model 1 and -0.940 in Model 2, indicating a
decrease in infant mortality during this time.

Demographic variables also exhibit notable associations. Maternal age is
negatively associated with infant mortality, with coefficients for age
groups 25–29 and 30–34 of -5.73 and -6.15 in Model 2, respectively. The
coefficient for Black mothers in Model 2 is 6.66, reflecting higher
infant mortality compared to other racial groups.

The models performed well, with low RMSE values and high $R^2$ values,
indicating that the predictors account for a substantial portion of the
variance in infant mortality rates.

# Discussion

## Key Findings and Implications

The analysis reveals distinct patterns in infant mortality rates across
various policy and demographic factors. In states where abortion
remained legal, the post-injunction period (after_injunction1) is linked
to reductions in infant mortality rates by 0.525 and 0.940 deaths per
1,000 live births in Models 1 and 2, respectively. However, states with
abortion bans (abortion_illegal1) exhibit significantly higher infant
mortality rates, with increases of 3.365 deaths in Model 1 and 1.400
deaths in Model 2. The DID interaction term (after_injunction1 ×
abortion_illegal1) reveals that abortion bans during the post-injunction
period result in an additional increase of 0.111 deaths in Model 1 and
0.280 deaths in Model 2. Demographic factors also play a critical role,
with older maternal age (25–34 years) reducing infant mortality rates,
while younger mothers (15–19 years) face elevated risks. Black mothers
experience disproportionately higher infant mortality rates, with an
increase of 6.660 deaths per 1,000 live births compared to Asian
mothers. These findings collectively emphasize the compounded effects of
policy and demographic disparities on infant health outcomes.

## Public Health Implications of Restrictive Abortion Laws

The findings highlight the significant public health consequences of
abortion bans in the United States. The observed increase of 0.285
infant deaths per 1,000 live births in states enforcing abortion bans
after the Dobbs decision underscores the detrimental impact of
restrictive abortion policies. These increases often reflect systemic
issues such as inadequate prenatal care, gaps in healthcare access, and
underlying socioeconomic inequities. States with pre-existing healthcare
challenges may experience amplified effects, suggesting an urgent need
for comprehensive policy interventions to address these disparities.
Tackling the root causes of these health disparities—such as improving
prenatal care, expanding healthcare access, and addressing socioeconomic
inequalities—will be essential for mitigating the unintended
consequences of restrictive abortion legislation.

## Racial and Socioeconomic Disparities

The results reveal stark racial disparities in infant mortality rates,
with Black mothers experiencing significantly higher rates compared to
other racial groups. These disparities are exacerbated by restrictive
abortion policies, which disproportionately affect already vulnerable
populations. The intersection of systemic healthcare inequities,
socioeconomic challenges, and restrictive legislation creates compounded
risks for marginalized communities, leading to worse health outcomes.
Addressing these inequities requires targeted policies that ensure
equitable access to maternal and infant healthcare, particularly in
communities most impacted by these disparities. Such policies should
prioritize investments in healthcare infrastructure, culturally
competent care, and programs that support at-risk populations.

## Integration of Policy and Healthcare Systems

The significant interaction between abortion legality and the
post-injunction period illustrates the compounded impact of restrictive
abortion policies on infant mortality rates. These findings underscore
the need for integrated healthcare systems that can address the
consequences of high-risk pregnancies, improve prenatal care, and ensure
equitable access to medical services. Policymakers must consider the
broader public health implications of abortion bans and implement
strategies that safeguard maternal and infant health. This includes
enhancing healthcare delivery in underserved areas, investing in
community-based interventions, and ensuring that restrictive policies
are accompanied by measures to mitigate their public health impact.

## Data and Temporal Limitations

A key limitation of our analysis is the temporal and demographic scope
of the dataset. The data includes only infant deaths under 1 year of age
occurring within the United States to U.S. residents during 2021–2022.
This timeframe does not capture the full year of infant mortality
outcomes following the Dobbs decision, as infants who died in 2023 are
not included. Furthermore, the dataset’s focus on infants under 1 year
of age may exclude cases that reflect longer-term health impacts. These
limitations could lead to an underestimation of the full effects of
abortion bans. Future research should incorporate more comprehensive and
updated datasets to provide a clearer and more complete analysis of the
long-term trends and implications of restrictive abortion policies.

## Weaknesses and Future Directions

While this study provides valuable findings, several limitations should
be addressed in future research. The analysis relies on aggregate
state-level data, which may obscure localized effects and intra-state
disparities. Unmeasured confounders, such as economic conditions,
healthcare infrastructure differences, or state-specific policies, could
also influence the results. Moreover, the short timeframe of the dataset
limits the ability to assess the long-term effects of abortion policies.
Future research should integrate more granular data, such as
county-level analyses, and extend the study period to capture
longer-term impacts. Qualitative studies focusing on the lived
experiences of affected populations can complement quantitative
analyses, offering a deeper understanding of the broader implications of
restrictive abortion laws. These efforts will support the development of
evidence-based policies aimed at reducing disparities and improving
maternal and infant health outcomes.

\newpage

\appendix

# Appendix {.unnumbered}

# Additional data details {#sec-data-details}

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive
check for the first model. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the
prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["First Model Posterior prediction check", "First Model Comparing the posterior with the prior","Second Model Posterior prediction check", "Second Model Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")
pp_check(second_model) +
  theme_classic() +
  theme(legend.position = "bottom")
posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
posterior_vs_prior(second_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
prior_summary(first_model)
prior_summary(second_model)
```

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-area
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["First Model Posterior prediction check", "First Model Comparing the posterior with the prior","Second Model Posterior prediction check", "Second Model Comparing the posterior with the prior"]

plot(
  first_model,
  "areas"
)
plot(
  second_model,
  "areas"
)
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")

plot(second_model, "trace")

plot(second_model, "rhat")
```

# Idealized Methodology for A Survey-Based Qualitative Studies {#sec-survey}

In our study, we examine the impact of abortion bans on infant mortality
rates in the United States following the Dobbs decision, employing both
observational data and survey-based research. Observational data provide
quantitative insights, revealing trends and associations across large
populations. However, they often fall short in addressing underlying
mechanisms due to limitations such as confounding factors and challenges
in establishing causality.

To address these gaps, in this appendix we propose incorporating surveys to gather
qualitative data, which provide rich, detailed accounts of individual
experiences and contextual factors. Qualitative methods allow for a
deeper understanding of how restrictive abortion policies affect
maternal health, access to healthcare, and infant outcomes, capturing
nuances that quantitative analysis alone cannot.

By combining quantitative and qualitative approaches, this study
utilizes the strengths of both methods. Quantitative data offer
generalizable findings and statistical rigor, while qualitative data
provide depth and context to interpret those findings. Together, they
create a more comprehensive framework for understanding the complex
relationship between restrictive abortion policies, maternal health, and
infant outcomes, enabling the development of nuanced, evidence-based
interventions and policy recommendations.

## Introduction

Restrictive abortion laws may exacerbate maternal stress, anxiety, or
depression—especially in cases of unintended or high-risk
pregnancies—potentially undermining maternal health and prenatal care.
Research indicates that maternal mental health is a critical determinant
of infant health outcomes, as conditions such as chronic stress,
depression, or anxiety during pregnancy can negatively impact prenatal
behaviors, access to healthcare, and overall maternal well-being. These
psychological challenges, compounded by societal pressures and systemic
barriers, can significantly increase the risk of adverse infant
outcomes, including infant mortality. The survey will focus on mothers
unable to access abortion services who subsequently experienced the
death of their infant, documenting their experiences to better
understand these interconnected factors.

To provide a comprehensive perspective, the study will also involve
healthcare providers and support organizations, capturing insights into
systemic challenges and the broader impacts of abortion bans on family
healthcare. By combining these perspectives with mothers’ lived
experiences, the survey aims to inform evidence-based interventions
addressing maternal psychological and healthcare needs and mitigating
the negative consequences of restrictive abortion policies.

This dual approach, integrating quantitative and qualitative methods,
seeks to uncover the complex relationship between restrictive abortion
laws, maternal mental health, and infant outcomes, forming the basis for
effective policy recommendations.

## Objective

The objective of this study is to investigate the psychological, social,
and systemic impacts of restrictive abortion laws on maternal health and
infant outcomes. By focusing on mothers unable to access abortion
services who experienced the death of their infant, the study aims to
document lived experiences, explore the role of maternal stress and
healthcare barriers in adverse outcomes, and gather perspectives from
healthcare providers and support organizations. Understanding how
maternal mental health directly affects prenatal care decisions, access
to healthcare, and overall maternal behavior is essential to addressing
the systemic factors contributing to higher infant mortality rates. The
findings will guide the development of evidence-based strategies to
improve maternal mental health, enhance healthcare access, and shape
policy responses to restrictive abortion laws.

## Sampling Approach

In this analysis, we will use non-probability Respondent-driven sampling to finding respondents. Respondent-driven sampling (RDS), developed by @heckathorn1997, is ideal for studying hidden populations lacking a sampling frame and facing potential stigma or harm. Unlike snowball sampling, RDS offers dual compensation for participation and recruitment, with recruitment occurring peer-to-peer via social networks. This method maintains confidentiality, reduces bias through statistical adjustments, and is particularly effective for sensitive topics, such as abortion, by leveraging trust within communities.

We use Respondent-Driven Sampling (RDS) for our survey because it is well-suited for reaching hidden or stigmatized populations, such as mothers who were unable to access abortion services and experienced the death of their infant. These individuals may be difficult to identify through conventional methods due to the sensitive nature of the topic and the lack of a clear sampling frame.

RDS utilizes peer networks, allowing participants to recruit others they trust, fostering higher participation rates in populations that may distrust external researchers. This approach helps maintain confidentiality and reduces the risk of social stigma. Additionally, RDS provides a structured framework for statistical adjustments, mitigating biases inherent in network-based recruitment, and improving the representativeness of the sample. These advantages make RDS an effective and ethical choice for gathering insights into the psychological, social, and systemic impacts of restrictive abortion laws.

## Target Population

Our target population is mothers who sought but were unable to access abortion services due to restrictive laws and experienced the death of their infant within the first year of life in states with abortion bans following the Dobbs decision.

## Sample frame

Given the sensitive nature of our study, identifying and reaching the initial participants poses a significant challenge. To address this, we aim to collaborate with support groups and advocacy organizations like Abortion Rights Advocacy Groups. These groups play a critical role in connecting researchers with individuals who are otherwise difficult to reach, particularly for sensitive topics like the impact of restrictive abortion laws on maternal and infant health.

Therefore, the sample frame of our study will be mothers who sought but were unable to access abortion services due to restrictive laws and experienced the death of their infant within the first year of life in states with abortion bans following the Dobbs decision and are reachable and can be contacted by Abortion Rights Advocacy Groups.

## Sample

We plan to survey 10 respondents for this study. Our sample will consist of mothers who fit the defined sample frame—those who sought but were unable to access abortion services due to restrictive laws, experienced the death of their infant within the first year of life, and are reachable through Abortion Rights Advocacy Groups. These mothers must also be willing to participate in and answer our survey questions, ensuring the reliability and depth of the collected data.

## Ethical Concerns

This study addresses sensitive and emotionally charged topics, necessitating a careful and ethical approach to protect participants’ well-being. We acknowledge the potential psychological distress that discussing experiences of unintended pregnancies, restrictive abortion laws, and infant loss may cause. To mitigate this, the survey will include trigger warnings, and participants will have the option to skip questions or withdraw at any time without penalty. All data collection will adhere to strict confidentiality protocols, ensuring that participants’ identities remain anonymous and their responses are securely stored. Recruitment through trusted advocacy groups will ensure that participants feel safe and supported throughout the process. Additionally, we will provide resources for mental health support to any participant who may experience discomfort as a result of participating in the study. This approach underscores our commitment to conducting ethical, respectful, and responsible research.

## Proposed Survey Design

\newpage

# References