### 2024-12-06 paper/ExaminingtheimpactOverturnofRoev.Wade:BanningofAbortiononInfantMortalityRatesintheUnitedStatesUsingaDifference-in-DifferencesApproach.qmd
update prose issue

---
title: "Impact of Roe v. Wade’s Overturn: Abortion Restrictions and Infant Mortality in the U.S. Using a Difference-in-Differences Analysis"
subtitle: "Infant death rates rose by 0.285 per 1,000 live births in abortion-restricted states compared to protected states."
author: 
  - Diana Shen
thanks: "Code and data are available at: [https://github.com/DianaShen1224/Relationship-between-infant-mortality-rate-and-prohibited-abortion](https://github.com/DianaShen1224/Relationship-between-infant-mortality-rate-and-prohibited-abortion)."
date: today
date-format: long
abstract: "This study investigates the impact of the Dobbs decision and abortion bans on U.S. infant mortality rates using 2021–2022 data. A Difference-in-Differences (DID) analysis shows a 0.285 increase in infant deaths per 1,000 live births in states enforcing bans, with amplified effects post-injunction. Maternal age and race are key factors, with Black mothers experiencing higher infant mortality rates. These findings highlight the public health impact of restrictive abortion policies and the need for targeted interventions to address disparities. This study uses the latest available data and rigorous modeling to explore the link between reproductive rights and infant health."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
editor: 
  markdown: 
    wrap: 72
---

```{r}
#| include: false
#| warning: false
#| message: false
library(tidyverse)
library(arrow)
library(knitr)
library(usmap)
library(ggplot2)
library(dplyr)
library(kableExtra)
library(rstanarm)
library(fixest)  
library(caret)
```

```{r}
#| include: false
#| warning: false
#| message: false
analysis_data<-read_parquet(here::here("data", "02-analysis_data", "analysis_data.parquet"))
```

# Introduction

The overturning of Roe v. Wade on June 24, 2022, through the Supreme
Court’s decision in Dobbs v. Jackson Women’s Health Organization, ended
nearly five decades of constitutional protection for abortion rights in
the United States. This landmark ruling shifted authority over abortion
regulations to individual states, creating a patchwork of laws
nationwide. The decision reignited debates about reproductive rights and
public health, highlighting disparities in access to abortion services
and their effects on health outcomes, especially in states with
restrictive policies.

Banning abortion usually has been linked to a rise in maternal mortality
rates, with Texas experiencing a 56% increase from 2019 to 2022
[@nbcnews]. However, restrictive abortion laws can also impact infant
mortality, as they often lead to more unintended or high-risk
pregnancies among disadvantaged populations with limited access to
healthcare. These pregnancies are associated with higher rates of
preterm births, low birth weights, and delivery complications, all of
which elevate infant mortality rates. Even before Roe v. Wade was
overturned, researchers highlighted the impacts of restrictive abortion
laws on mortality rates. @Burdick2024 found that states with 11–12
restrictive laws during 2014–2018 had a 16% higher infant mortality
rate, particularly in counties with higher proportions of Black
populations, inadequate prenatal care, and maternal smoking. Similarly,
@Harper2023 showed that restrictive laws are linked to higher maternal,
fetal, and infant mortality, states with moderate abortion legislation
had significantly lower maternal mortality (25.79 per 100,000 live
births) and infant mortality (20.56 per 1,000 live births) compared to
restrictive states, underscoring the health benefits of less restrictive
policies.

The estimand of our analysis is the change in the difference in infant
death rates between states that banned abortion and states where
abortion remained legal after June 24, 2022. The object of the
estimation is to estimate the change in the difference in the infant
death rate between the treatment and control groups based on
observational data in the United States.

Instead of examining the impact of abortion laws over a broader time
periods like prior research, our study focuses on the effect of abortion
bans on infant mortality rates in the United States specifically from
2021 to 2022, with an emphasis on changes following the 2022 Dobbs
decision. Using a Difference-in-Differences (DID) approach, we assess
the interaction between abortion legality in states and the timing of
the injunction implemented to understand its influence on infant health
outcomes. The study utilizes observational data from the Centers for
Disease Control and Prevention’s WONDER database [@cdc_wonder_2024], captures infant
mortality trends across states with varying levels of abortion
restrictiveness. Key predictors include the mother’s state of residence,
race, age, and the year of infant death, enabling a detailed examination
of the effects of restrictive abortion laws on public health.

Our Results show that the overturning of Roe v. Wade and the
implementation of abortion bans are linked to a rise in infant mortality
rates in states where abortion became illegal. Through a
Difference-in-Differences (DID) approach, we estimate that states
enforcing abortion bans after the injunction saw an increase of 0.285
deaths per 1,000 live births compared to states where abortion remained
legal. Furthermore, we find that maternal age and race are key
determinants of infant mortality, with older maternal age groups
associated with lower mortality rates and Black mothers experiencing
significantly higher rates. These results emphasize the public health
consequences of restrictive abortion laws and the urgent need for
policies that mitigate these disparities.

This study is important as we highlight the public health implications
of abortion bans, particularly their role in increasing infant mortality
rates. By focusing on the 2022 Dobbs decision, we provide timely
evidence on how restrictive abortion policies disproportionately affect
vulnerable populations, emphasizing the need for evidence-based policies
to mitigate these adverse outcomes.

The structure of this paper is as follows: @sec-data details the data
sources and variables used in the analysis. @sec-model outlines the
modeling approach, including assumptions and specifications of the
Bayesian DID framework. In @sec-results, we present the findings,
emphasizing key observations. In @sec-results, we present findings on
the impact of abortion bans and the post-injunction period, highlighting
their association with increased infant mortality. Key disparities by
maternal age and race are discussed, along with model validation metrics
like RMSE and $R^2$. Finally, @sec-discussion explores the implications
of the results and suggests potential directions for future research.
Additional details are provided in the appendices: [Appendix-
@sec-data-details] elaborates on external data sources, [Appendix-
@sec-model-details] provides an in-depth explanation of the model, and
[Appendix- @sec-survey] introduces the idealized methodology for related
qualitative analysis.

# Data {#sec-data}

## Overview

We conduct our infant death rate observational data analysis using the R
programming language [@citeR]. Our dataset, obtained from the CDC
Wonder[@cdc_wonder_2024], based on data as of December 2022, provides detailed state-level information on infant death rates in the United States, serving as the foundation for our study of the impact of abortion injunctions. CDC WONDER
(Wide-ranging ONline Data for Epidemiologic Research) is an online
system providing public health professionals and the public with access
to a wide range of the Centers for Disease Control and Prevention (CDC)
epidemiologic data and resources. This dataset offers detailed state-level information on infant death rates in the United States, serving as the foundation for our study of the impact of abortion injunctions.
Following the guidelines outlined in
@tellingstories, we analyze the impact of abortion injunctions on infant
death rates using the Difference-in-Difference method. This approach, shown in @sec-model, utilizes the main effects and the interaction between the legality of abortion in the mother’s state of residence and the timing of the infant’s death as the predictor variable. Additionally, we include the infants' year of death, race, and age as control variables to account for potential confounding factors, and the mother’s state of residence as a random variable.

We selected the CDC WONDER dataset for its combination of granularity, consistency, and accessibility, making it suitable for a nationwide analysis. Alternative datasets lacked either the breadth or the standardization required for a study of this scale. Additionally, CDC WONDER’s public availability enhances transparency and allows for independent verification, ensuring the reliability and reproducibility of our research.

In this study, we utilized several R packages to streamline data
manipulation, modeling, and visualization. The `tidyverse` package
[@tidyverse] was central to our workflow, providing tools for data
wrangling and analysis, while `arrow` [@arrow] efficiently managed
parquet files for larger datasets. We used `modelsummary` package
[@modelsummary] to generate clean and interpretable summaries of our
regression models. We used `knitr` [@knitr] and `kableExtra`
[@kableExtra] to generate clean tables and reproducible reports, and
`usmap` for geographical visualizations. We use `dplyr` [@dplyr] to manipulate data. We use `testthat` [@testthat] to test the simulated and analysis dataset. We use `usmap` [@usmap] to plot and analyze U.S. state and county-level data on maps by providing preformatted map geometries and easy integration with data. We use `here` [@here] package simplifies the management of file paths, ensuring a reproducible
workflow by creating a consistent point of reference for file locations,
regardless of the operating system or working directory. Data cleaning
was facilitated by `janitor` [@janitor], while `lubridate` [@lubridate]
handled date-time variables. Visualization tasks were enhanced with
`ggplot2` [@ggplot2], and Bayesian regression modeling was conducted
using `rstanarm` [@rstanarm]. We use `fixest` [@fixest] to conduct the
event study. We use `caret` [@caret] to conduct the cross-validation of
the model. Together, these packages supported an efficient, reproducible
workflow for analyzing and presenting our findings.

## Measurement

The process of translating real-world events into our dataset requires a
systematic approach to measurement and data gathering. In this research,
we analyze infant mortality rates in the United States to understand the
impact of abortion policies and demographic factors. The CDC’s Linked Birth/Infant Death Records database provides detailed data, including
maternal and infant characteristics, to support this study. Variables
such as maternal age, race, education, birth weight, gestational age,
and prenatal care are included to ensure a detailed examination of
factors influencing infant health outcomes.

The data, sourced from birth and death certificates, undergoes a
thorough cleaning and validation process to address inconsistencies and
missing information, ensuring accuracy. Infant mortality rates are
calculated per 1,000 live births, with additional indicators such as
reliability labels for rates based on fewer than 20 deaths. To further
enhance data accuracy, weights are applied to adjust for unlinked infant
death records, ensuring that all deaths—linked and unlinked—are
proportionally represented. Weights are calculated for each state and
infant age-at-death cohort (\<7 days, 7-27 days, and 28 days to 1 year)
using the formula:
$\text{Weight} = \frac{\text{Linked deaths} + \text{Unlinked deaths}}{\text{Linked deaths}}$
This adjustment means areas with complete linkages have a weight of 1.0,
while areas with incomplete linkages are adjusted proportionally.
Additionally, imputation techniques are employed to handle missing
birthweight data, assigning values based on gestational period, race,
sex, and plurality, further reducing potential bias in the analysis.

Once cleaned, the data is stratified by maternal and infant attributes,
providing an understanding of trends and disparities across demographic
groups. Statistical models are employed to estimate associations between
state-level abortion policies and infant mortality, controlling for
confounders such as maternal characteristics. This structured approach
transforms complex datasets into actionable understanding, highlighting
on the health implications of varying abortion laws across the U.S. By
utilizing detailed datasets and robust statistical techniques, this
study offers a detailed understanding of the intersection between
policy, demographics, and infant health outcomes, informing the public
health strategies and legislative decisions.

## Outcome variables

### Death Rate

The death rate (`death rate`) refers to the number of deaths of infants under one
year of age per 1,000 live births in a specified population during a
given time period. It serves as a key indicator of a population’s
overall health and access to healthcare. Rates labeled as "Unreliable"
occur when there are fewer than 20 deaths in the numerator, as these
figures do not meet the National Center for Health Statistics (NCHS)
standards for reliability or precision. After cleaning the dataset, the "Unreliable" label is converted into a separate variable named `reliable` (This variable is available in the `analysis data` for reference but was not included in the model analysis.).

```{r}
#| label: fig-deathrate
#| fig-cap: >
#|   Distribution of Infant Death Rates: The infant death rate (per 1,000 live births) 
#|   shows a peak between 5 and 10 deaths per 1,000 infants, with a right-skewed 
#|   distribution. 
#| echo: false

ggplot(analysis_data, aes(x = death_rate)) +
  geom_histogram(binwidth = 1, fill = "brown", color = "black") +
  labs(
    x = "Infant Death Rate (per 1,000 live births)", y = "Count",
    caption = "Data source: CDC Wonder."
  ) +theme_classic()
```

@fig-deathrate illustrates the distribution of infant death rates (per
1,000 live births) based on the available data. The majority of infant
death rates cluster around lower values, with a peak observed between 5
and 10 deaths per 1,000 infants. As the death rate increases, the
frequency of occurrences decreases, indicating that higher infant death
rates are relatively rare. There are few instances where the rate
exceeds 20 deaths per 1,000 infants, suggesting these are outliers.
Overall, the distribution is right-skewed, highlighting the predominance
of lower infant death rates in the dataset.

## Predictor variables

### Abortion Legality

Abortion legality (`abortion_illegal`) is a newly
created variable in our dataset, which refers to the legal status of
abortion within a given state. It is represented as a binary variable,
where 1 indicates that abortion is illegal in the state, and 0 indicates
that abortion is legal. This classification reflects the policies and
laws governing abortion access in each state.

@fig-map illustrates the legality of abortion across U.S. states. Based
on @reproductiveright, after June 2022, abortion has been illegal in
Alabama, Arkansas, Idaho, Indiana, Kentucky, Louisiana, Mississippi,
Missouri, North Dakota, Oklahoma, South Dakota, Tennessee, Texas, West
Virginia, and Wisconsin following the June 2022 decision.

```{r}
#| include: false
#| warning: false
#| message: false

# Ensure the number of states and `abortion_illegal` values match
illegal_states <- tolower(c("Alabama", "Arkansas", "Idaho", "Indiana", "Kentucky", 
                    "Louisiana", "Mississippi", "Oklahoma", "South Dakota", 
                    "Tennessee", "Texas", "West Virginia"))
abortion_status <- data.frame(
  state = tolower(c(
    "Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado", "Connecticut", 
    "Delaware", "Florida", "Georgia", "Hawaii", "Idaho", "Illinois", "Indiana", 
    "Iowa", "Kansas", "Kentucky", "Louisiana", "Maine", "Maryland", "Massachusetts", 
    "Michigan", "Minnesota", "Mississippi", "Missouri", "Montana", "Nebraska", 
    "Nevada", "New Hampshire", "New Jersey", "New Mexico", "New York", "North Carolina", 
    "North Dakota", "Ohio", "Oklahoma", "Oregon", "Pennsylvania", "Rhode Island", 
    "South Carolina", "South Dakota", "Tennessee", "Texas", "Utah", "Vermont", 
    "Virginia", "Washington", "West Virginia", "Wisconsin", "Wyoming"
  )),
  short = c(
    "AL", "AK", "AZ", "AR", "CA", "CO", "CT", "DE", "FL", "GA", "HI", "ID", "IL", "IN", 
    "IA", "KS", "KY", "LA", "ME", "MD", "MA", "MI", "MN", "MS", "MO", "MT", "NE", 
    "NV", "NH", "NJ", "NM", "NY", "NC", "ND", "OH", "OK", "OR", "PA", "RI", "SC", 
    "SD", "TN", "TX", "UT", "VT", "VA", "WA", "WV", "WI", "WY"
  )
)|>mutate(abortion_illegal = as.factor(case_when(state %in% illegal_states ~ 1, TRUE ~ 0)))
```

```{r}
#| include: false
#| warning: false
#| message: false
us_map <- map_data("state")
map_data <- us_map|>
   left_join(abortion_status, c("region"="state"))|>filter(!is.na(abortion_illegal))
```

```{r}
#| label: fig-map
#| fig-cap: >
#|   Abortion Legality by State: After June 2022, abortion is prohibited in the
#|   following states: Alabama, Arkansas, Idaho, Indiana, Kentucky, Louisiana,
#|   Mississippi, Missouri, North Dakota, Oklahoma, South Dakota, Tennessee,
#|   Texas, West Virginia, and Wisconsin.
#| echo: false
#| fig-pos: "H"
ggplot() +
  geom_polygon(data = map_data, aes(fill = abortion_illegal, x = long, y = lat, group = group),
    color = "black") +
  scale_fill_manual(
    values = c("0" = "blue", "1" = "red"),
    labels = c("Legal", "Illegal"),
    name = "State is Illegal for Abortion or Not"
  )+
  theme_void() +
  labs(
    caption = "Source: Center for Reproductive Rights\nDate as of Nov.19 2024"
  ) +
  theme(legend.position = "bottom")
```

### After Injunction

Another created variable After Injunction (`after_injunction`) indicates whether the date
of infants' death falls after the implementation of the abortion
injunctions following the overturn of Roe v. Wade in June 2022. It is
coded as a binary variable, where 1 represents dates after June 2022,
signifying the post-injunction period, and 0 represents dates before
June 2022, indicating the pre-injunction period.

@fig-rateandstate illustrates trends in infant death rates (per 1,000
live births) over time in states where abortion is illegal versus legal.
In states with abortion bans, infant mortality shows a slight increase
after June 2022, marked by the dashed line representing the overturn of
Roe v. Wade. In contrast, states where abortion remains legal exhibit
relatively stable trends. This suggests a potential link between
abortion bans and rising infant mortality rates, warranting further
investigation.

```{r}
#| label: fig-rateandstate
#| fig-cap: >
#|   Infant Death Rates Over Time: Infant death rates (per 1,000 live births)
#|   show a slight increase in states where abortion became illegal following
#|   June 2022, whereas rates in states with legal abortion remained relatively
#|   stable. 
#| echo: false
#| message: false
#| fig-pos: "H"
analysis_data|>
  mutate(illegal = if_else( abortion_illegal== 1, "Illegal State", "Legal State"))|>
  ggplot(aes(x = date, y = death_rate,colour = illegal)) +
  geom_point(alpha = 0.5)+
  geom_smooth() +
  labs(x = "Date of Infant Death", y = "Infant Death Rate \n(per 1,000 live births)",color="State is Illegal for Abortion or Not",caption = "Data source: CDC Wonder.\nIn June 2022, the U.S. Supreme Court overturned Roe v. Wade.") +
  facet_wrap(vars(illegal), nrow = 2)+
  geom_vline(xintercept = as.Date("2022-06-01"), linetype = "dashed")  +
  scale_x_date(
    limits = c(as.Date("2021-01-01"), as.Date("2022-12-01")), 
    date_breaks = "3 months",  
    date_labels = "%Y-%m" 
  )+
  scale_color_brewer(palette = "Set1")+
  theme_classic() +
  theme(legend.position = "bottom")
```

### Interaction Term: Abortion Legality and After Injunction

To evaluate the impact of abortion injunctions on infant death rates, we
utilize the Difference-in-Difference (DID) method. As part of this
approach, we construct an interaction term defined as the product of
`abortion_illegal` and `after_injunction`. The resulting interaction
term equals 1 when abortion is illegal the time is after the injunction,
and 0 otherwise. This term is included in the analysis to evaluate the
joint impact of these factors on infant death rates.

```{r}
#| label: tbl-interact
#| tbl-cap: >
#|   Average Infant Death Rate by Abortion Legality and Injunction Timing: 
#|   States categorized as "Illegal Abortion and After Injunction" ('1') 
#|   exhibit a higher average infant death rate (per 1,000 live births) compared 
#|   to states in other categories, including "Legal Abortion and Before/After 
#|   Injunction" or "Illegal Abortion and Before Injunction" ('0'). 
#| echo: false
#| warning: false
#| message: false

analysis_data_state <- analysis_data|>mutate(illegal_after_injunction=as.numeric(as.character(abortion_illegal))*as.numeric(as.character(after_injunction)))|>
  group_by(illegal_after_injunction)|>rename(`Abortion Restriction Post-Overturn`=illegal_after_injunction)|>
  summarize(`Average Infant Death Rate (per 1,000 live births)` = round(mean(death_rate, na.rm = TRUE), 1))

kable(analysis_data_state, align = "c") |> 
  footnote(
    "Data Source: CDC Wonder"
  )
```

@tbl-interact highlights average infant death rates (per 1,000 live
births) based on abortion legality and injunction timing. States
categorized as ‘1’, where abortion became illegal after the injunction,
show a higher average infant death rate of 8.3 compared to 7.9 in ‘0’
states, where abortion remained legal or abortion used to be legal
before the injunction. This suggests that restrictive abortion policies
and their enforcement may contribute to higher infant mortality,
potentially due to increased unintended or high-risk pregnancies,
reduced access to maternal healthcare, and exacerbated social or
economic disparities.

## Control Variables

Our analysis includes three key control variables: maternal age
(`age_of_mother`), race (`mothers_single_race`), and year of death
(`year_of_death`), to account for demographic and temporal differences.

```{r}
#| label: tbl-summary
#| tbl-cap: >
#|   Demographic Distribution of Infant Mortality Cases by Year, Maternal Age, and Race (2021–2022): 
#|   The data highlights the proportional representation of cases across different years, 
#|   maternal age groups, and racial categories. Younger mothers and Black or African American 
#|   mothers show higher proportions of cases compared to other groups.
#| echo: false
#| warning: false
#| message: false
analysis_data|>mutate(year_of_death=as.factor(year_of_death))|>
  select(year_of_death,age_of_mother,mothers_single_race) |> 
  rename(`Year of Death` = year_of_death,
         `Age of Mother` = age_of_mother,
         `Mother's Single Race` = mothers_single_race)|>
  modelsummary::datasummary_skim()
```

@tbl-summary shows the summary statistics of control variables. 50.7% of infant deaths occurred in 2022, slightly higher than in 2021 (49.3%). The majority of cases were among
mothers aged 25–29 (37.8%), followed by those aged 30–34 (32.5%), and
20–24 (27.2%), with the lowest occurrence in mothers aged 15–19 (2.6%).
By race, White mothers accounted for the highest percentage (75.7%),
followed by Black or African American mothers (24.0%), and Asian mothers
(0.3%).

### Year of Death

The Year of Death (`year_of_death`) represents the calendar year in which the
infant death occurred, specifically focusing on the years 2021 and 2022
in this study. This variable is essential for analyzing temporal trends
and identifying patterns or shifts in infant mortality rates over this
essential period. By controlling the year of death, we can closely
examine the impact of significant policy changes, such as the
overturning of Roe v. Wade in 2022, as well as other temporal factors on
infant health outcomes.

### Age of Mother

The Age of the Mother (`age_of_mother`) refers to the age group of the
mother at the time of childbirth. In our observational data, age is
categorized into the brackets 15–19 years, 20–24 years, 25–29 years, and
30–34 years in the raw data, in compliance with privacy requirements.

@tbl-motherage displays the average infant death rate (per 1,000 live
births) by the age group of mothers. It shows a clear trend: younger
mothers, particularly those aged 15–19 and 20–24, have higher infant
death rates (9.0 and 10.0, respectively), while older mothers in the
25–29 and 30–34 age groups have lower rates (7.4 and 6.6, respectively).
This suggests that maternal age may be a factor influencing infant
mortality, with younger mothers experiencing higher risks.

```{r}
#| label: tbl-motherage
#| tbl-cap: >
#|   Average Infant Death Rate by Maternal Age Group: 
#|   Infant mortality rates (per 1,000 live births) vary significantly by maternal age, 
#|   with younger mothers experiencing higher rates. Mothers aged 20–24 exhibit the 
#|   highest infant death rate at 10.0, while the lowest rate of 6.6 is observed among 
#|   mothers aged 30–34.
#| echo: false
#| warning: false
#| message: false

analysis_data_state <- analysis_data|>
  group_by(age_of_mother)|>
  summarize(`Average Infant Death Rate (per 1,000 live births)` = round(mean(death_rate, na.rm = TRUE), 1)) |>rename(`Age of Mother`=age_of_mother)

# Display the result as a table
kable(analysis_data_state)|>footnote("Data Source: CDC Wonder")
```

### Mother's Single Race

Mother's Single Race (`mothers_single_race`) refers to the self-identified race of the mother, which was recorded as a single race category. In this dataset, the available categories are Asian, Black or African American, and White. This classification facilitates the analysis of maternal and infant health outcomes across these racial groups.

```{r}
#| label: tbl-motherrace
#| tbl-cap: >
#|   Average Infant Death Rate by Mother’s Single Race: 
#|   Infant mortality rates (per 1,000 live births) vary significantly by race, 
#|   with Black or African American mothers experiencing the highest rate at 14.5, 
#|   compared to 5.8 for White mothers and 5.4 for Asian mothers. 
#| echo: false
#| warning: false
#| message: false
analysis_data_state <- analysis_data|>
  group_by(mothers_single_race)|>
  summarize(`Average Infant Death Rate (per 1,000 live births)` = round(mean(death_rate, na.rm = TRUE), 1)) |>rename(`Mother's Single Race`=mothers_single_race)

# Display the result as a table
kable(analysis_data_state)|>footnote("Data Source: CDC Wonder")
```

@tbl-motherrace represents the average infant death rates (per 1,000 live births) by maternal race. Black or African American mothers have the highest rate at 14.5, compared to 5.8 for White mothers and 5.4 for Asian mothers. These differences indicate significant disparities in infant mortality outcomes across racial groups.

## Random Variable

### State

State (`state`) referes to the state of the mother’s legal residence at the time of birth. This
data is categorized by the mother’s state of residence and excludes territories.

```{r}
#| label: fig-choromap
#| fig-cap: >
#|   Infant Death Rates Across U.S. States (per 1,000 Live Births): 
#|   The infant mortality rate is significantly higher in the Southeastern states, 
#|   with Alabama and Mississippi reporting the highest rates. In contrast, 
#|   states like California and New York exhibit notably lower rates.
#| echo: false
#| fig-pos: "H"
death<-analysis_data|>group_by(state)|>summarise(death_rate=mean(death_rate))|>mutate(state=tolower(state))
death_full <- us_map|>
   full_join(death, c("region"="state"))

ggplot() +
  # First layer: Fill with death_rate
  geom_polygon(
    data = death_full,
    aes(x = long, y = lat, group = group, fill = death_rate),
    color = "black"
  ) +
  scale_fill_viridis_c(
    name = "Infant Death Rate\n(per 1,000 live births)",
    option = "C",
    na.value = "grey50",
    breaks = c(10, 15, 20,25),  # Custom breaks for the color legend
  ) +
  theme_void() +
  labs(
    caption = "Grey indicates missing data\nSource: Center for Reproductive Rights\nDate as of December, 2022"
  ) +
  theme(legend.position = "bottom")
```

@fig-choromap is a choropleth map that visualizes the average infant
death rate per 1,000 live births across U.S. states, with data
categorized by color intensity. States with higher infant death rates
are shown in lighter shades (yellow), while states with lower rates are
represented in darker shades (purple). States with missing data are
indicated in grey. The map highlights disparities in infant mortality
rates, with some states, particularly in the southeastern U.S.,
exhibiting notably higher rates compared to others. States like Alabama
and Mississippi show the highest rates, while states such as California
and New York exhibit lower rates.

# Model {#sec-model}

For our analysis, we use a Bayesian Difference-in-Differences (DID)
model to quantify the impact of abortion injunctions on infant death
rate across states in the USA. This method compares changes in infant
death rates over time between a treatment group (states where abortion
injunctions were implemented) and a control group (states where abortion
remains legal), providing a robust framework for estimating causal
effects.

## Difference-In-Difference Approach (DID)

Difference-in-Differences (DID) is a causal inference method that
estimates the impact of an intervention by comparing changes in outcomes
over time between a treatment group and a control group. By assuming
parallel trends in the absence of the intervention, the DID framework
isolates the intervention’s effect through **the interaction between
time periods (pre- and post-intervention) and group membership**. This
makes it a robust method for evaluating policy changes and other
significant events.

### Assumption

The validity of the Difference-in-Differences (DID) approach in our
study relies on several key assumptions, which, if violated, could
affect the robustness and reliability of our results [@cunningham2021]:

#### Parallel Trends Assumption

The model relies on the assumption that, without the abortion
injunction, the treatment group (states where abortion became illegal)
and the control group (states where abortion remained legal) would have
followed similar trends in infant mortality rates. However, this
assumption can be challenged by differences in healthcare systems,
socioeconomic conditions, or pre-existing policy environments, which may
cause deviations from parallel trends.

#### Compositional Differences

In addition to parallel trends, compositional changes in repeated
cross-sectional data over time can influence results. For instance,
shifts in demographic factors such as maternal age or racial makeup
between pre-and post-intervention periods may confound the observed
changes in infant mortality rates. Including control variables helps
adjust for these confounders, ensuring that the estimated effect
reflects the policy intervention rather than changes in the population
composition.

#### Long-term Effects vs. Reliability

The length of the analysis period further complicates the reliability of
the results. While longer periods allow for capturing long-term effects,
they also increase the likelihood of unrelated factors, such as new
policies or healthcare advancements, influencing the outcome. Over time,
some states in the control group may also implement similar abortion
policies, diluting the treatment effect. Conversely, shorter analysis
periods

#### Functional Form Dependence

Finally, the model assumes that the functional relationship between
predictors and outcomes is correctly specified. If this assumption is
violated, differences in outcomes between treatment and control groups
may stem from misspecified relationships rather than the intervention
itself. This highlights the importance of carefully assessing the
model’s structure and assumptions to ensure valid results.

## Model Parameters

The model focuses on the infant death rate per 1,000 live births as the
dependent variable and includes the following independent variables:

-   **Treatment Group Indicator (`abortion_illegal`)**: A binary variable
    indicating the group: 1 if abortion is illegal in the state. 0 if
    abortion remains legal.

-   **Time Indicator (`after_injunction`)**: A binary variable indicating
    the time period: 0 for the period before the injunction was
    implemented. 1 for the period after the injunction was implemented.

-   **Interaction Term (`after_injunction` × `abortion_illegal`)**: The
    interaction term captures the additional change in the outcome
    (e.g., percentage of support, behavior, or health-related metrics)
    for regions where abortion is illegal after the injunction
    implemented, beyond any changes observed in regions where abortion
    remains legal.

Additionally, we include these control variables to adjust for
demographic and temporal factors that could influence infant mortality
rates, ensuring that the estimated effects are not confounded by these
characteristics:

-   **Age of Mother (`age_of_mother`)**: The age of the mother at the time of birth,
    represented as a continuous variable.

-   **Mother's Single Race (`mothers_single_race`)**: Self-identified race of the mother,
    represented as a categorical variable.

-   **Year of Death (`year_of_death`)**: The year in which the infant death occurred,
    represented as a continuous variable.

Finally, we include a random variable:

-   **State (`state`)**: Mother's residence state when the infant was born as a
    random effect in the model.

By isolating the impact of the abortion injunction through this
interaction term `after_injunction`×`abortion_illegal`, our model
identifies changes in infant death rates that can be attributed to the
policy intervention. Additional details, validation of model assumption,
and diagnostics supporting the model’s implementation is provided in
[Appendix- @sec-model-details].

## Model set-up

In our analysis, we employ two models. The first model includes only the
predictor variables and the random effect for the state. The second
model builds on the first by incorporating additional control variables
to account for potential confounding factors.

Define $y_i$ as as the infant death rate per 1,000 live births, and
$\gamma_j$ as the random state effect for state $j$. 
$$
\begin{aligned}
y_i|\mu_i,\sigma &\sim \mbox{Normal}(\mu_i, \sigma)\\
\text{First Model}: \mu_i &= \beta_0 + \beta_1\cdot\text{After Injunction} + \beta_2\cdot\text{Abortion Illegal}\\&+\beta_3\cdot\text{After Injunction}\cdot\text{Abortion Illegal}_i+\gamma_j\\
\text{Second Model}:\mu_i &= \beta_0 + \beta_1\cdot\text{After Injunction}_i + \beta_2\cdot\text{Abortion Illegal}_i\\&+ \beta_3\cdot\text{After Injunction}_i\cdot\text{Abortion Illegal}_i\\&+\beta_4\cdot\text{Year of Death}_i+\beta_5\cdot\text{Age of Mother}_i+\beta_6\cdot\text{Mother's Single Race}_i+\gamma_j\\
\beta_0 &\sim \mbox{Normal}(0, 2.5)\\
\beta_1 &\sim \mbox{Normal}(0, 2.5)\\
\beta_2 &\sim \mbox{Normal}(0, 2.5)\\
\beta_3 &\sim \mbox{Normal}(0, 2.5)\\
\beta_4 &\sim \mbox{Normal}(0, 2.5)\\
\beta_5 &\sim \mbox{Normal}(0, 2.5)\\
\beta_6 &\sim \mbox{Normal}(0, 2.5)\\
\gamma_j&\sim \text{Normal}(0, \sigma_j^2)\\
\sigma &\sim \mbox{Exponential}(1)
\end{aligned}
$$

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm` used for both Gaussian Bayesian models.

### Model justification

#### Analysis Framework

For our analysis, we employ Bayesian regression models to estimate
infant death rates in states where abortion injunctions were implemented
and states where abortion remains legal. The causal effect of the
abortion injunction is measured using the interaction term between
after_injunction and abortion_illegal. To ensure consistency and
comparability, both models share the same predictors, capturing the
determinants of infant mortality.

One model includes control variables — Age of Mother,
Mother's Single Race, and Year of Death to adjust for confounders
that may influence infant death rates. This adjustment addresses
compositional differences assumed in the DID framework and ensures
unbiased estimates of the intervention’s effect. The second model
excludes these controls to focus on the overall intervention impact,
avoid overfitting, and prioritize simplicity where variability is
adequately accounted for by design or random effects. By fitting two
models, we compare outcomes to refine estimates and evaluate the direct
impact of the intervention.

#### Model Design and Rationale

Our models include state as a random effect to account for regional
variations in infant mortality that might introduce bias. Using a random
effect rather than a fixed effect captures variability across states
without estimating individual effects, modeling unobserved
heterogeneity—such as differences in healthcare systems and
socioeconomic conditions—through a single variance term. This approach
avoids overfitting, enhances efficiency, and ensures generalizability.
By adjusting for unmeasured state-specific factors, the models strike a
balance between parsimony and interpretability, making them well-suited
for our analysis.

#### Why DID Over Alternative Models?

We selected the Difference-in-Differences (DID) framework because
randomization is infeasible given the pre-existing differences between
states in healthcare systems, socioeconomic conditions, and cultural
factors. DID utilizes time and group-level variations, controlling for
unobserved, time-invariant confounders. The parallel trends assumption
ensures that untreated states provide a valid counterfactual for treated
states. Interaction terms, such as after_injunction and
abortion_illegal, allow for direct measurement of policy effects, while
random effects adjust for regional variability.

Alternative models, such as Fixed Effects, Propensity Score Matching
(PSM), and Interrupted Time Series (ITS), were considered but deemed
less suitable. Fixed Effects cannot estimate the effects of variables
that do not vary over time within a unit. PSM balances observed
covariates but does not account for temporal dynamics. ITS analyzes
trends within a single group but lacks a comparison group. DID’s ability
to integrate temporal and group-level differences makes it the most
robust and appropriate framework for estimating the impact of abortion
bans on infant mortality.

#### Bayesian Framework

We use a Bayesian approach to enhance the Difference-in-Differences
(DID) method, which allows us to incorporate prior knowledge and better
account for uncertainty. With the `rstanarm` package, we apply weakly
informative priors—such as a Normal prior centered at zero with moderate
spread ($\mu=0$, $\sigma=2.5$) for fixed effects and an Exponential
prior (rate = 1) for variability parameters. By using `autoscale=TRUE`, these priors automatically adjust to the scale of the data, improving model stability and ensuring they complement rather than dominate the observed data. This approach
helps us center estimates realistically, account for hidden differences
between groups, and produce reliable results while assuming trends
between treated and control groups are similar before treatment.

#### Limitations and Applicability

Our Bayesian DID model offers a robust approach to examining the impact
of abortion bans on infant mortality but has limitations. Misspecified
priors, data imbalances, and violations of the parallel trends
assumption could introduce biases. Despite these challenges, the model
effectively integrates prior knowledge and addresses uncertainty,
providing an important understanding of policy impacts.

# Results {#sec-results}

## Model Results

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false
first_model <-
  readRDS(file = here::here("models/first_model.rds"))
second_model <-
  readRDS(file = here::here("models/second_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Summary of Bayesian Regression Model Examining the Impact of Abortion Bans on Infant Mortality Rates in the United States: A Difference-in-Differences Approach Using 2021-2022 Infant Mortality Data"
#| warning: false

modelsummary::msummary(
  list(first_model,second_model),
  fmt = 3,coef_omit="state",notes = "Note. All models include a random variable for the state. The reference level of age_of_mother is 15-19. The reference level of monthers_single_race is Asian. Source: CDC Wonder")
```

The results from the two Bayesian models are summarized in
@tbl-modelresults. The interaction term between after_injunction and
abortion_illegal is positive and significant (0.111 in
first model and 0.280 in the second model), indicating a change in infant
mortality rates associated with the combination of abortion bans and the
post-injunction period.

For the independent effects, states where abortion is illegal show
higher infant mortality rates, with coefficients of 3.365 in the first
model and 1.400 in the second model. The post-injunction period alone is
associated with coefficients of -0.525 in the first model and -0.940 in
the second model, indicating a decrease in infant mortality during this
time.

Demographic variables also exhibit some associations. Maternal age is
negatively associated with infant mortality, with coefficients for age
groups 25–29 and 30–34 of -5.73 and -6.15 in the second model, relative
to groups 15-19 respectively. The coefficient for Black mothers in
the second model is 6.66 relative to Asia mothers, reflecting a higher
infant mortality compared to other racial groups.

The model highlights key demographic and policy-driven influences on
infant mortality rates. The interaction between abortion bans and the
post-injunction period is associated with increased mortality, while
states with abortion bans consistently exhibit higher rates. Younger
mothers and Black mothers face significantly elevated risks,
highlighting disparities likely driven by limited healthcare access and
systemic inequities. Conversely, increasing maternal age correlates with
lower infant mortality rates, reflecting improved access to resources,
better prenatal care, and greater maternal stability. These results
demonstrate the impact of policy and demographic factors on infant
health outcomes.

Both models exhibit low RMSE values and high $R^2$, indicating a strong overall fit and that the predictors explain much of the variance in infant mortality rates. The second model, with its additional covariates, provides a more detailed and adjusted view of the intervention’s effects, while the first model offers a simpler and more direct perspective. The choice between models depends on the balance between interpretability and precision.

# Discussion {#sec-discussion}

## Key Findings and Implications

The analysis explains notable patterns in infant mortality rates
influenced by policy and demographic factors. For states where abortion
remained legal, the post-injunction period (after_injunction1) is
associated with reductions in infant mortality rates, with a decrease of
0.525 deaths per 1,000 live births in the first model and 0.947 in the
second model, compared to the pre-injunction period. In contrast, states
where abortion became illegal (abortion_illegal1) experienced
significantly higher infant mortality rates, with increases of 3.365
deaths in the first model and 1.401 in the second model, relative to
states where abortion remained legal.

The Difference-in-Differences (DID) interaction term (after_injunction1
× abortion_illegal1) indicates that abortion bans during the
post-injunction period resulted in an additional increase of 0.111
deaths per 1,000 live births in the first model and 0.280 in the second
model. These effects highlight the compounded impact of restrictive
abortion policies on infant health outcomes.

Demographic factors also play a significant role. Compared to mothers
aged 15–19 years (reference level), older maternal age groups (20–24,
25–29, and 30–34 years) are associated with progressively lower infant
mortality rates, with reductions of 3.571, 5.732, and 6.155 deaths per
1,000 live births in the second model, respectively. Regarding maternal
race, Black or African American mothers have a significantly higher
infant mortality rate, with an increase of 6.604 deaths per 1,000 live
births compared to Asian mothers (reference level). In contrast, White
mothers exhibit a lower infant mortality rate, with a decrease of 2.562
deaths per 1,000 live births compared to Asian mothers.

These findings underscore the interplay of policy and demographic
factors, emphasizing the disproportionate burden of restrictive abortion
policies and systemic inequities on infant mortality rates.

## Impact of Control Variables on Model Performance and Results

The inclusion of control variables in the second model significantly improves its explanatory power compared to the first model, as evidenced by the higher adjusted $R^2$ (0.778) in the second model versus 0.213 in the first. These control variables—such as maternal age, race, and year of death—capture detailed demographic and temporal variations influencing infant mortality rates, leading to more precise and reliable estimates of the impact of abortion bans.

For instance, the coefficient for the interaction term (`after_injunction` × `abortion_illegal`) increases from 0.111 in the first model to 0.285 in the second model, suggesting a stronger observed effect of abortion injunctions after adjusting for confounding factors. Additionally, the second model highlights disparities, such as the substantial positive effect of 6.604 for Black mothers on infant mortality, underscoring the influence of racial and demographic factors.

The improvement in model fit is further demonstrated by a lower RMSE (2.36 in the second model versus 4.21 in the first) and better predictive performance metrics, including LOOIC and WAIC. These results emphasize the importance of controlling for demographic and socioeconomic heterogeneity across states to ensure that the estimated effects of abortion bans are robust and not confounded by underlying differences.

## Racial and Socioeconomic Disparities

The results explain stark racial disparities in infant mortality rates,
with Black mothers experiencing significantly higher rates compared to
other racial groups. These disparities are exacerbated by restrictive
abortion policies, which disproportionately affect already vulnerable
populations. The intersection of systemic healthcare inequities,
socioeconomic challenges, and restrictive legislation creates compounded
risks for marginalized communities, leading to worse health outcomes.
Addressing these inequities requires targeted policies that ensure
equitable access to maternal and infant healthcare, particularly in
communities most impacted by these disparities. Such policies should
prioritize investments in healthcare infrastructure, culturally
competent care, and programs that support at-risk populations.

## Addressing Public Health and Long-Term Implications of Restrictive Abortion Laws

Restrictive abortion policies significantly impact maternal and infant
health, as evidenced by an increase of 0.285 infant deaths per 1,000
live births in states enforcing abortion bans during the post-injunction
period. These policies exacerbate systemic issues such as inadequate
prenatal care, gaps in healthcare access, and socioeconomic inequities,
disproportionately affecting marginalized communities. Over time, the
consequences may include higher rates of untreated maternal mental
health issues, delayed or insufficient prenatal care, and elevated
infant mortality, creating lasting health challenges for families and
communities. Addressing these disparities requires targeted
interventions, such as expanding maternal mental health services,
improving prenatal and postnatal care in underserved areas, ensuring
culturally competent healthcare, and advocating for equitable policies.
Investments in community-based support systems, enhanced healthcare
delivery, and robust data infrastructure are essential to monitor and
mitigate the long-term impacts of these policies, fostering sustainable
improvements in maternal and infant health outcomes.

## Data and Temporal Limitations

A key limitation of our analysis is the temporal and demographic scope
of the dataset. The data includes only infant deaths under 1 year of age
occurring within the United States to U.S. residents during 2021–2022.
This timeframe does not capture the full year of infant mortality
outcomes following the Dobbs decision, as infants who died in 2023 are
not included. Furthermore, the dataset’s focus on infants under 1 year
of age may exclude cases that reflect longer-term health impacts. These
limitations could lead to an underestimation of the full effects of
abortion bans. As no publicly available dataset currently provides
complete 2023 data or longer-term health outcomes, future research
should incorporate more detailed and updated datasets to provide a
clearer and more complete analysis of the long-term trends and
implications of restrictive abortion policies.

## Weaknesses and Future Directions

While this study provides meaningful findings, several limitations
should be addressed in future research. The analysis relies on aggregate
state-level data, which may obscure localized effects and intra-state
disparities. Unmeasured confounders, such as economic conditions,
healthcare infrastructure differences, or state-specific policies, could
also influence the results. Moreover, the short timeframe of the dataset
limits the ability to assess the long-term effects of abortion policies.
Future research should integrate more granular data, such as
county-level analyses, and extend the study period to capture
longer-term impacts. Qualitative studies focusing on the lived
experiences of affected populations can complement quantitative
analyses, offering a deeper understanding of the broader implications of
restrictive abortion laws. These efforts will support the development of
evidence-based policies aimed at reducing disparities and improving
maternal and infant health outcomes.

\newpage

\appendix

# Appendix {.unnumbered}

# Additional data details {#sec-data-details}

## Raw Dataset Download

The complete process of downloading the raw dataset is shown in the `README.md` file.

## Dataset and Graph Sketches

Sketches depicting both the desired dataset and the tables generated in
this analysis is available in the GitHub Repository `other/sketches`.

## Data Cleaning

The raw dataset, consisting of U.S. infant mortality records from 2021
and 2022, was cleaned and pre-processed to ensure its suitability for
analysis. The dataset was first imported from a tab-separated file, with
irregular lines handled appropriately. Variables were standardized using
`janitor::clean_names()` to ensure consistent naming conventions. Key
variables such as state, maternal age, maternal race, year of death,
month, and infant death rates were selected for analysis, and missing
values were removed.

Additional variables were created to facilitate analysis:
`after_injunction`, indicating whether the data correspond to the period
after June 2022 (post-Dobbs decision), and `abortion_illegal`, denoting
whether the state had restrictive abortion laws. Death rate reliability
was assessed, and unreliable rates were flagged for transparency. The
death rate column was cleaned to remove annotations and converted to
numeric format for analysis. A date variable was created by combining
the year and month columns, providing a timeline for analysis. Finally,
a date variable was created by combining the year and month variables.
The cleaned dataset was saved in CSV and Parquet formats to ensure
compatibility with analytical workflows.

## Attribution Statement

The data utilized in this study was sourced from the CDC WONDER Online
Database. Access to the data and its use complies with the terms
outlined in the [CDC WONDER Data Use
Agreement](https://wonder.cdc.gov/datause.html). Specifically, the data
was used for academic research purposes, with acknowledgment of the
Centers for Disease Control and Prevention (CDC) as the original data
provider. The CDC, however, does not assume responsibility for any
analyses, interpretations, or conclusions drawn from the data by the
authors of this study.

# Model details {#sec-model-details}

## Model Card

The model card, referenced from @ozoani2022modelcard, is included in the README under the section titled `Model Card for Infant Mortality and Abortion Policy Analysis Model` for detailed documentation.

## Parall Assumption: An Event Study Analysis

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false

cleaned_data <- analysis_data %>%filter(year_of_death==2022)|>
  mutate(relative_time = (month_code-6))

# Round relative_time to account for fractional months if necessary
cleaned_data <- cleaned_data %>%
  mutate(relative_time = as.factor(round(relative_time, 0)))|>mutate(abortion_illegal=as.numeric(as.character(abortion_illegal)))

est_did <- feols(
  death_rate ~ mothers_single_race + age_of_mother + i(relative_time , abortion_illegal, ref = -1) | state + relative_time, 
  data = cleaned_data
)
```

```{r}
#| label: fig-eventstudy
#| fig-cap: >
#|   Event Study Analysis of the Impact of Abortion Bans on Infant Mortality Rates in the United   States: 
#|   Evidence Using June 2022 as the Treatment Date and Pre-Treatment (-1, May 2022) as the Reference Period.
#| warning: false
#| echo: false
#| eval: true
#| fig-pos: "H"

iplot(est_did,main = "Impact of Abortion Bans on Infant \nMortality Rates by State",
  value.lab = "Estimate and __ci__ Conf. Int.",xlab = "Relative Time (Month)")
```

@fig-eventstudy depicts the estimated effects of abortion bans on infant
mortality rates over time, with relative months on the x-axis and
coefficient estimates with 95% confidence intervals on the y-axis.
Before the implementation of abortion bans ($t<0$), the coefficients are
centered around zero, indicating no significant differences in infant
mortality rates between treated and control states, supporting the
parallel trends assumption. Post-ban ($t\geq0$), the estimates show a
slight upward trend, suggesting a potential increase in infant mortality rates in restricted states; however, the wide confidence intervals, which overlap with zero, indicate that these effects are not statistically significant. Overall, the results do not provide
conclusive evidence of a causal relationship between abortion bans and
changes in infant mortality rates, emphasizing the need for further
analysis with more precise data or extended time frames.

## Posterior predictive check

In @fig-ppcheck-1 and @fig-ppcheck-2 we implement a posterior predictive
check for the first model and second model. It represents posterior
predictive checks for the first model and second model, illustrating how
well each model captures the observed data. The posterior predictive
checks for the first model (policy-focused) and the second model (policy and
demographic factors) illustrate model fit. In @fig-ppcheck-1, the
observed ($y$) and replicated ($y_{rep}$) distributions align
moderately, reflecting general trends but limited precision.
@fig-ppcheck-2 shows improved alignment, indicating that adding
demographic factors enhances predictive accuracy and better captures the
complexities of infant mortality rates.

In @fig-posteriorvsprior1 and @fig-posteriorvsprior2 we compare the
posterior with the prior. The posterior vs. prior plots for the first model and second model illustrate the role of data in informing the posterior estimates of the model parameters. In the first model
@fig-posteriorvsprior1, which focuses on policy-related predictors (e.g., abortion restrictions), the data significantly shifts the
posterior distributions for key variables, demonstrating their
explanatory power. The second model @fig-posteriorvsprior2, incorporating additional demographic factors such as maternal age and race, shows a more refined and concentrated posterior distribution, emphasizing the
added predictive utility of demographic variables. The comparison
indicates that the second model provides better-informed parameter
estimates, supporting the hypothesis that demographic factors play an
essential role in understanding infant mortality in the context of
restrictive abortion laws.

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheck
#| layout-ncol: 1
#| fig-cap: "Posterior Predictive Check: Examining how the model for infant death rate fits, and is affected by, the data"
#| fig-subcap: ["First Model Posterior prediction check", "Second Model Posterior prediction check" ]
#| fig-pos: "H"

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")
pp_check(second_model) +
  theme_classic() +
  theme(legend.position = "bottom")
```

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-posteriorvsprior1
#| fig-cap: "Comparing Prior and Posterior Distributions: Examining how the first model for infant death rate fits, and is affected by, the data"
#| fig-pos: "H"
posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-posteriorvsprior2
#| fig-cap: "Comparing Prior and Posterior Distributions: Examining how the second model for infant death rate fits, and is affected by, the data"
#| fig-pos: "H"
posterior_vs_prior(second_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

```{r}
#| eval: false
#| echo: false
#| message: false
#| warning: false
#| include: false

prior_summary(first_model)
prior_summary(second_model)
```

## Prior Summary for Models

| **Model**        | **Component**     | **Specified Prior**                              | **Adjusted Prior**                               | **Description**                                         |
|--------------|--------------|--------------|-------------------|------------------|
| **First Model**  | Intercept         | Normal(0, 2.5)                                   | Normal(0, 12)                                    | Prior for intercept after centering predictors.         |
|                  | Coefficients      | Normal(\[0,0,0\], \[2.5,2.5,2.5\])               | Normal(\[0,0,0\], \[26.84,28.43,45.17\])         | Prior for coefficients, adjusted for predictor scaling. |
|                  | Auxiliary (sigma) | Exponential(1)                                   | Exponential(0.2)                                 | Prior for the standard deviation of residuals.          |
|                  | Covariance        | Decov(reg. = 1, conc. = 1, shape = 1, scale = 1) | Decov(reg. = 1, conc. = 1, shape = 1, scale = 1) | Regularizing prior for covariance.                      |
| **Second Model** | Intercept         | Normal(0, 2.5)                                   | Normal(0, 12)                                    | Prior for intercept after centering predictors.         |
|                  | Coefficients      | Normal(\[0,0,0,...\], \[2.5,2.5,2.5,...\])       | Normal(\[0,0,0,...\], \[26.84,28.43,24.78,...\]) | Prior for coefficients, adjusted for predictor scaling. |
|                  | Auxiliary (sigma) | Exponential(1)                                   | Exponential(0.2)                                 | Prior for the standard deviation of residuals.          |
|                  | Covariance        | Decov(reg. = 1, conc. = 1, shape = 1, scale = 1) | Decov(reg. = 1, conc. = 1, shape = 1, scale = 1) | Regularizing prior for covariance.                      |

: Prior Specifications for Bayesian Models {#tbl-priorsummary}

@tbl-priorsummary shows the prior summaries for the two models, indicating that both use weakly
informative priors. For the intercept, a normal prior with a mean of 0
and a standard deviation of 2.5 was specified, which adjusts to a scale
of 12 after centering the predictors. The coefficients for both models
are also assigned normal priors with a mean of 0 and a scale of 2.5,
which are adjusted based on the predictors, resulting in scales such as
\[26.84, 28.43, 45.17\] for the first model and \[26.84, 28.43,
24.78,...\] for the second model. The auxiliary parameter (sigma)
follows an exponential prior with a rate of 1, adjusted to 0.2, while a
decov prior regularizes the covariance structure in both models.
Although the prior structure is similar, the second model incorporates
more predictors, leading to a larger number of coefficients with
slightly adjusted scales. These priors are designed to be weakly
informative, allowing the data to primarily guide the posterior
estimates while avoiding extreme parameter values.

## Distribution

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-area
#| layout-ncol: 2
#| fig-cap: "95% credibility intervals for fitted models"
#| fig-subcap: ["First Model", "Second Model"]
#| fig-pos: "H"

plot(
  first_model,
  "areas"
)

plot(
  second_model,
  "areas"
)
```

Credibility intervals are the Bayesian equivalent of confidence
intervals, offering an understanding of the range of probable values for
a parameter [@tellingstories]. Using Bayesian estimation, we derive a
posterior distribution for each model coefficient. @fig-area-1 presents
the 95% credibility intervals for coefficients in the first model, while
@fig-area-2 illustrates the same for the second model. These intervals
represent the range within which 95% of the posterior probability mass
falls, offering a robust understanding of parameter uncertainty and the
influence of predictors under the Bayesian framework [@tellingstories].

## Cross Validation

```{r}
#| echo: false
#| eval: false
#| message: false
#| warning: false
#| include: false

set.seed(123)
# Split the data into training and testing sets (80/20 split)
train_index <- createDataPartition(analysis_data$death_rate, p = 0.8, list = FALSE)
train_data <- analysis_data[train_index, ]
test_data <- analysis_data[-train_index, ]

# Fit a Bayesian model on the training set
model1 <- stan_glmer(
  death_rate ~ after_injunction * abortion_illegal+ (1 | state),
  data = train_data,
  family = gaussian(),
  prior = normal(0, 2.5),
  prior_intercept = normal(0, 2.5),
  prior_aux = exponential(1),
  iter = 2000, chains = 4, seed = 123
)
# Make predictions on the test set
predictions <- posterior_predict(model1, newdata = test_data)
predicted_means <- rowMeans(predictions)  # Average over posterior samples

# Calculate RMSE manually
errors <- test_data$death_rate - predicted_means
rmse_value <- sqrt(mean(errors^2))

# Output the RMSE
print(paste("The RMSE is:", round(rmse_value, 2)))
```

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| include: false

set.seed(123)
train_index <- createDataPartition(analysis_data$death_rate, p = 0.8, list = FALSE)
train_data <- analysis_data[train_index, ]
test_data <- analysis_data[-train_index, ]
model2 <- stan_glmer(
  death_rate ~ after_injunction * abortion_illegal + year_of_death + age_of_mother + mothers_single_race + (1 | state),
  data = train_data,
  family = gaussian(),
  prior = normal(0, 2.5),
  prior_intercept = normal(0, 2.5),
  prior_aux = exponential(1),
  iter = 2000, chains = 4, seed = 123
)
# Make predictions on the test set
predictions <- posterior_predict(model2, newdata = test_data)
predicted_means <- rowMeans(predictions)  # Average over posterior samples

# Calculate RMSE manually
errors <- test_data$death_rate - predicted_means
rmse_value <- sqrt(mean(errors^2))

# Output the RMSE
print(paste("The RMSE is:", round(rmse_value, 2)))
```

The RMSE values from cross-validation indicate that the first model and the second model perform similarly in predicting infant mortality rates,
with RMSE values of 5.07 and 5.06, respectively. This similarity
suggests that adding demographic factors such as maternal age and race
in the second model has only a marginal impact on predictive accuracy
compared to the first model, which primarily relies on policy-related
variables like abortion legality and timing of the injunction. An average error of approximately 5 deaths per 1,000 live births indicates
the magnitude of prediction errors, which may or may not be acceptable
depending on the variability in the dataset. While the second model
slightly reduces the error, its added complexity may not justify the
trade-off if simplicity is a priority. However, if a detailed understanding of demographic influences is essential, the second model
provides a modest improvement in accuracy while offering a more detailed understanding of maternal and infant health disparities. The
cross-validation process reinforces the reliability of these models by
confirming their performance across multiple data splits, reducing the
risk of overfitting.

## Diagnostics

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay1
#| fig-cap: "First Model: Checking the convergence of the MCMC algorithm for the infant death rate model"
#| fig-subcap: ["First Model Trace plot", "First Model Rhat plot"]
#| layout-ncol: 1
#| fig-pos: "H"

plot(first_model, "trace")

plot(first_model, "rhat")
```

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay2
#| fig-cap: "Second Model: Checking the convergence of the MCMC algorithm for the infant death rate model"
#| fig-subcap: ["Second Model Trace plot", "Second Model Rhat plot"]
#| layout-ncol: 1
#| fig-pos: "H"
plot(second_model, "trace")

plot(second_model, "rhat")
```

The trace plot in @fig-stanareyouokay1-1 shows the sampling chains for
the first model, indicating good mixing and convergence as the chains
overlap and appear stable throughout. This suggests that the model has
achieved convergence during sampling. The Rhat plot in
@fig-stanareyouokay1-2 demonstrates that all Rhat values are below 1.05,
further confirming convergence and indicating that the model’s posterior
samples are reliable. @fig-stanareyouokay2-1 represents the trace plots
for the second model, showing the posterior samples across four chains.
The plots indicate good mixing, as the chains overlap well and stabilize
around the same region, suggesting convergence. @fig-stanareyouokay2-2 is
the Rhat diagnostic plot for the second model. All Rhat values are close
to 1.00, confirming that the Markov chains have converged, and the
posterior estimates are reliable. This supports the validity of the
second model’s Bayesian estimation process.

# Idealized Methodology for A Survey-Based Qualitative Studies {#sec-survey}

Our study examines the impact of abortion bans on infant mortality rates
in the U.S. following the Dobbs decision, combining observational
Difference-in-Differences (DID) analysis with qualitative surveys. While the DID analysis quantifies policy impacts, qualitative surveys provide
a context-specific understanding of how restrictive abortion policies
affect maternal mental health, healthcare access, and pregnancy
decisions. These lived experiences help interpret quantitative findings,
offering a fuller understanding of the systemic and psychological
factors driving observed trends, and supporting evidence-based policy
recommendations.

## Introduction

Restrictive abortion laws can heighten maternal stress, anxiety, or
depression—particularly in unintended or high-risk
pregnancies—compromising maternal health and prenatal care. Research from
@crugnola2022 highlights maternal mental health as a key determinant of
infant health outcomes, with conditions like chronic stress and depression adversely affecting prenatal behaviors, healthcare access,
and overall well-being. These compounded challenges, driven by societal and systemic barriers, can significantly increase risks of adverse
infant outcomes, including mortality.

Our survey focuses on mothers unable to access abortion services who
later experienced the death of their infant, documenting their
experiences to better understand these interconnected factors. By
integrating quantitative data with qualitative perspectives, this study
aims to inform evidence-based interventions, addressing maternal
psychological and healthcare needs, and mitigating the adverse effects
of restrictive abortion policies.

## Objective

The objective of this study is to investigate the psychological, social,
and systemic impacts of restrictive abortion laws on maternal health and
infant outcomes. By focusing on mothers unable to access abortion
services who experienced the death of their infant, the study aims to
document lived experiences and explore the role of maternal stress and
healthcare barriers in adverse outcomes. Understanding how maternal
mental health directly affects prenatal care decisions, access to healthcare, and overall maternal behavior is essential to addressing the
systemic factors contributing to higher infant mortality rates. The
findings will guide the development of evidence-based strategies to improve maternal mental health, enhance healthcare access, and shape
policy responses to restrictive abortion laws.

## Sampling Approach

In this analysis, we use non-probability Respondent-Driven Sampling
(RDS), developed by @heckathorn1997, is a recruitment method designed
for hidden populations lacking a sampling frame, where participants
recruit others through their social networks, with dual compensation for
participation and recruitment, ensuring confidentiality and mitigating
biases. RDS enables peer-to-peer recruitment with dual compensation for
participation and recruitment, fostering trust within communities. This method maintains confidentiality, reduces bias through statistical adjustments, and is particularly effective for sensitive topics like abortion.

We use Respondent-Driven Sampling (RDS) for our survey as it is
effective for accessing hidden or stigmatized populations, such as
mothers unable to access abortion services and who experienced the death
of their infant. RDS utilizes peer networks for recruitment, fostering
trust and participation in sensitive contexts while maintaining
confidentiality and reducing social stigma. Its statistical adjustments
mitigate biases in network-based recruitment, enhancing sample
representativeness. These strengths make RDS an ethical and effective
method for examining the psychological, social, and systemic impacts of
restrictive abortion laws.

## Target Population

Our target population is mothers who sought but were unable to access
abortion services due to restrictive laws and experienced the death of
their infant within the first year of life in states with abortion bans
following the Dobbs decision.

## Sample frame

Given the sensitive nature of this study, identifying and reaching
participants presents a significant challenge. To address this, we will
collaborate with support groups and advocacy organizations, such as Abortion Rights Advocacy Groups, which can connect researchers with individuals otherwise difficult to reach. These organizations are
essential for engaging with populations affected by restrictive abortion
laws, particularly concerning maternal and infant health.

Thus, the sample frame for our study comprises mothers who reside in
areas where abortion services are inaccessible due to restrictive laws,
have experienced the death of their infant within the first year of
life, and are reachable through advocacy organizations following the
Dobbs decision.

## Sample

We aim to survey 20 respondents who meet our defined sample criteria:
mothers unable to access abortion services due to restrictive laws, who
experienced the death of their infant within the first year of life, and
who are reachable through Abortion Rights Advocacy Groups. Participation
will be voluntary, and respondents must be willing to answer the survey
questions, ensuring the quality and depth of the data collected.

## Recruitment of Respondents

To recruit participants for this sensitive study, we will employ a
Respondent-Driven Sampling (RDS) approach in collaboration with Abortion
Rights Advocacy Groups, leveraging their established connections with
our target population. This organization will act as initial "seeds" to
identify mothers affected by restrictive abortion laws who meet our
study criteria. Participants may also invite others from their network
with similar experiences, allowing the sample to grow while maintaining
confidentiality. Outreach materials emphasizing anonymity and the
study’s significance will be shared through trusted channels. Interested
participants will undergo a screening process and provide informed
consent, with the option to complete the survey online for safety and
privacy. To ensure participant well-being, mental health resources will
be offered throughout the process, fostering trust and encouraging
meaningful engagement.

## Handling Non-response bias

Non-response bias is a concern as it can distort survey results and lead
to unrepresentative conclusions about the target population. Since our
survey addresses a sensitive topic and takes approximately 10–15 minutes
to complete, there is a risk of participant dropout or refusal. To
minimize this, we provide a clear overview of the survey’s purpose,
confidentiality measures, and estimated completion time, fostering trust
and encouraging participation.

## Respondent Validation

To ensure data credibility, we implement a robust respondent validation
process. Participants will first undergo screening to confirm
eligibility, including inability to access abortion services due to
restrictive laws and experiencing infant loss within the first year of
life. Responses will be assessed for completeness and consistency, and
duplicate entries will be removed to maintain integrity. Partnering with
trusted support groups and advocacy organizations enhances reliability,
ensuring the findings accurately represent the target population’s
experiences.

## Ethical Concerns

This study addresses sensitive and emotionally charged topics, requiring
an ethical approach to safeguard participants’ well-being. Recognizing
the potential psychological distress associated with discussing
unintended pregnancies, restrictive abortion laws, and infant loss, the
survey will include trigger warnings and allow participants to skip
questions or withdraw at any time without penalty. Strict
confidentiality protocols will protect participants’ identities, and
responses will be securely stored. Recruitment through trusted advocacy
groups will foster a safe and supportive environment, while mental
health resources will be provided to those who experience discomfort
during the study. This framework reflects our commitment to ethical,
respectful, and responsible research practices.

## Proposed Survey Design

Abortion, or even considering abortion, carries significant social
stigma. Therefore, our survey requires extremely careful phrasing and
design to minimize this stigma’s impact on the study of abortion,
pregnancy, fertility, and related demographic and health outcomes. In
@lindberg2022, researchers tested innovative methods in a national
survey to address persistent underreporting of abortion in the U.S.,
focusing on question design and placement to reduce sensitivity, social
desirability bias, and perceived intrusiveness. Although none of their
experimental approaches significantly improved abortion reporting
compared to the control condition, their study offers meaningful
guidelines for improving abortion measurement. We aim to adapt our
survey design based on their methodologies to enhance data accuracy and
respondent comfort.

This survey explores the impacts of restrictive abortion laws on
maternal mental health and infant outcomes, emphasizing neutral,
carefully crafted questions to ensure participant comfort and minimize
social desirability bias. Inspired by @lindberg2022, it employs
strategies like thoughtful question placement, phrasing, and opt-out
options (e.g., "Prefer not to say") to address underreporting in
sensitive topics. Combining multiple-choice and open-ended formats, the
survey aims to balance structured data collection with detailed understanding, fostering trust and enhancing data accuracy. Also, based
on @lindberg2022 findings that respondents preferred introductions
emphasizing how sharing their abortion experiences could help improve
health services for other women, we adopted a "helping" framework in
each section of our survey introductions to encourage accurate reporting
and increase participants’ motivation to contribute.

## Solution to the response bias in our survey

We draw on recommendations from @stantcheva2023 to minimize response
biases. Common response biases identified in survey design include
moderacy bias, extreme response bias, ordering bias, acquiescence bias,
experimenter demand effect(EDE), and social desirability bias (SDB). Our
survey primarily focuses on strategies to reduce moderacy bias, extreme
response bias, ordering bias, and SDB. The detailed definitions of these
biases are provided in [Appendix- @sec-definition].

To mitigate bias, we enhance our survey in the following ways, drawing
on recommendations from @stantcheva2023:

-   Addressing Extreme/Moderacy Bias: We use a minimum of five response
    options for scale questions to provide more detailed choices,
    reducing the likelihood of respondents defaulting to extreme or
    middle answers.

-   Mitigating Response Order Bias: For nominal questions, we randomize
    response options.

-   Minimizing Social Desirability Bias (SDB): The survey design
    addresses social desirability bias (SDB) by emphasizing anonymity
    and confidentiality throughout. A clear introduction outlines the
    survey’s purpose—academic research on the impact of restrictive
    abortion laws—and reassures participants that their responses will
    remain confidential and solely used for research. The anonymous
    online format creates a safe environment for participants to share
    their experiences without fear of judgment or stigma. A feedback
    section at the end encourages participants to express concerns or share additional thoughts, fostering trust and enhancing data
    quality.

### Survey Link

The survey has been implemented using Google Forms. You can access it
here: [Survey Link](https://forms.gle/YRrb6Z6EyorHPkuF8).

## Copy of Survey on Restrictive Abortion Laws and Maternal Health

Welcome Section

Introduction: Welcome to our study on the impact of restrictive abortion
laws on maternal and infant health. Your participation in this survey
will help us understand the psychological, social, and systemic impacts
of these laws. Rest assured that your responses are anonymous and will
only be used for academic research purposes.

This survey is conducted by nonpartisan researchers in public health and
social sciences. It consists of 17 carefully designed questions and
should take approximately 10–15 minutes to complete.

Please answer the questions honestly. If you experience any discomfort
while completing the survey, you may stop at any time. For support, we
provide a list of mental health resources at the end of the survey.

Contact Information: Diana Shen Email: diana.shen\@mail.utoronto.ca

Section 1: Demographics and Background Information

1.  What is your age?
    -   Under 18
    -   18–24
    -   25–34
    -   35–44
    -   45–54
    -   55+
    -   Prefer not to say
2.  What is your highest level of education?
    -   Less than high school
    -   High school graduate or equivalent
    -   Some college
    -   Bachelor’s degree
    -   Graduate or professional degree
    -   Prefer not to say
3.  What is your marital status? 
    - Single 
    - Married 
    - Divorced 
    - Widowed 
    - Prefer not to say

Section 2: Understanding Abortion Experiences

Introduction: This section focuses on understanding abortion
experiences, including the circumstances and decisions surrounding them.
Your responses are invaluable in helping researchers and policymakers
improve health services and support for women and families. Please know
that your answers are entirely confidential and will only be used for
research purposes. If you are comfortable, we encourage you to answer as
honestly as possible. If you prefer not to answer, you are welcome to
skip this section.

1.  Are you willing to answer this part?
    -   Yes
    -   No \[Jump to Section 3\]
2.  Did you seek abortion services during your pregnancy?
    -   Yes
    -   No
    -   Prefer not to say
3.  If yes, were you unable to access abortion services due to legal
    restrictions in your state?
    -   Yes
    -   No
    -   Prefer not to say
4.  How did the inability to access abortion services impact your mental
    health during pregnancy?
    -   No impact
    -   Mild impact
    -   Moderate impact
    -   Severe impact
    -   Prefer not to say
5.  Did you receive any support from healthcare providers or community
    organizations during your pregnancy?
    -   Yes
    -   No
6.  If you were unable to access abortion services, what barriers did
    you encounter? (Select all that apply)
    -   Legal restrictions
    -   Financial constraints
    -   Lack of healthcare providers
    -   Distance to clinic
    -   Fear of stigma or judgment
    -   Prefer not to say
7.  If you sought an abortion but were unable to access one, how did
    this affect your mental health during pregnancy?
    -   Increased stress
    -   Anxiety
    -   Depression
    -   Anger or frustration
    -   Feeling of helplessness
    -   Prefer not to say

Section 3: Experiences and Support for Women Facing Pregnancy Challenges

Introduction: This section seeks to understand the experiences and
outcomes of women who, due to restrictive abortion laws, were unable to
access abortion services and subsequently faced the loss of their child.
We recognize that discussing past pregnancies can be sensitive,
especially those involving circumstances such as abortion or the loss of
a child. Your understanding is invaluable in identifying areas where
healthcare and support services can be improved. If you prefer not to
answer, you are welcome to skip this section.

1.  Are you willing to answer this part?
    -   Yes
    -   No \[Jump to Section 4\]
2.  Did you experience the loss of your child within the first year of
    life?
    -   Yes
    -   No \[Jump to Section 4\]
    -   Prefer not to say
3.  What was the primary cause of your child’s death as communicated by
    healthcare providers?
    -   Premature birth or related complications
    -   Congenital abnormalities or genetic conditions
    -   Sudden Infant Death Syndrome (SIDS)
    -   Infections (e.g., pneumonia, sepsis)
    -   Birth trauma or delivery complications
    -   Lack of access to timely medical care
    -   Other (please specify)
4.  To what extent do you believe the instances during your pregnancy
    where mental health challenges prevented you from seeking or
    receiving adequate medical care?
    -   1 (No impact)
    -   2
    -   3
    -   4
    -   5 (Significantly)
5.  To what extent do you believe the mental health challenges caused by
    the inability to access abortion services contributed to health
    complications for your child?
    -   Not at all – I don’t believe my mental health challenges had any
        impact on my child’s health.
    -   Slightly – I think there may have been a minor impact on my
        child’s health.
    -   Moderately – I feel my mental health challenges had a noticeable
        impact on my child’s health.
    -   Significantly – I believe my mental health challenges had a
        considerable impact on my child’s health.
    -   Completely – I think my mental health challenges were the
        primary factor in my child’s health complications.
    -   Prefer not to say
6.  Looking back, do you believe access to abortion services could have
    positively affected your mental and physical health during
    pregnancy?
    -   Yes
    -   No
    -   Maybe
    -   Prefer not to say

Section 4: Perspectives and Support

If you feel distressed or need support after completing this survey, the
following resources are available to provide assistance:

1.  Postpartum Support International (PSI)
    -   Website: www.postpartum.net
    -   Helpline: 1-800-944-4773 (Text "Help" to 800-944-4773)
    -   Services: Support for mental health during and after pregnancy,
        including peer support and counseling.
2.  Mental Health America (MHA)
    -   Website: www.mhanational.org
    -   Services: Online screening tools, support networks, and
        educational materials.
3.  SAMHSA National Helpline
    -   Website: www.samhsa.gov
    -   Hotline: 1-800-662-HELP (4357)
    -   Services: Free, confidential referrals for mental health and
        substance use disorders.

Section 5: Feedback

1.  Do you have any concerns or feedback regarding the survey or entity?
    -   Your feedback is important to us and will help ensure
        transparency and trust in the research process.

Section 6: Thank You

Thank you for taking the time to complete this survey. Your honest
feedback is invaluable and will help us better understand and address
the experiences of women who have faced similar circumstances. We deeply
appreciate your participation and the courage it takes to share your
experiences.

## Response bias definition {#sec-definition}

In the design of the questionnaire, there will be some common biases
that may occur when running the questionnaire.

@stantcheva2023 defines these biases as:

-   **Moderacy response bias**: The tendency to choose mid-scale
    responses.

-   **Extreme response bias**: The inclination to select extreme values
    on the scale.

-   **Response order bias**: Occurs when the position of options
    influences responses, with the **primacy effect** (favoring early
    options in written surveys) and **recency effect** (favoring later
    options in oral surveys) as common forms.

-   **Social desirability bias**: Respondents may hide true views to
    present a socially favorable image, influenced by topic and social
    context.

-   **Acquiescence bias**: The habit of giving positive responses, such
    as consistently choosing "agree" or "yes." \newpage

# References